%
% nats-relay.tex
%
% Z Specification for Biff NATS Relay
% Formal model of connection lifecycle, KV watcher state machine,
% and session presence management.
%

\documentclass[a4paper,10pt,fleqn]{article}
\usepackage[margin=1in]{geometry}
\usepackage{fuzz}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\begin{document}

\title{Biff NATS Relay: A Z Specification}
\author{Formal Model of Connection Lifecycle and KV Watcher}
\date{March 2026}
\hypersetup{
  pdftitle={Biff NATS Relay: A Z Specification},
  pdfauthor={punt-labs},
  pdfsubject={Z Specification},
  pdfcreator={fuzz/probcli}
}
\maketitle

\tableofcontents
\newpage

%%----------------------------------------------------------------------
\section{Introduction}
%%----------------------------------------------------------------------

The NATS relay is the networked backend for biff's communication
system.  It manages a TCP connection to a NATS server, lazily
provisions infrastructure (KV bucket, JetStream streams), and runs
a KV watcher that detects session changes and wall broadcasts in
real-time.

This specification models three interacting state machines:

\begin{enumerate}
  \item \textbf{Connection lifecycle} --- lazy connect, disconnect,
    reconnect, and close transitions.
  \item \textbf{KV watcher} --- the long-lived watcher that observes
    session and wall changes, survives snapshot-done markers and
    timeouts, and restarts only on errors.
  \item \textbf{Session presence} --- KV-backed sessions with TTL
    expiry, heartbeat refresh, and logout detection.
\end{enumerate}

The key invariant proven by this model: the KV watcher remains
alive across snapshot-done markers and idle timeouts.  Only
explicit shutdown or unrecoverable errors terminate it.  This
captures the fix for biff-udp.

A second invariant, added to close a recurring class of wedge
bugs, concerns the connection itself.  A NATS connection can
enter a \emph{half-open} state: the TCP socket is up
(\texttt{is\_closed} is false) but every JetStream or KV request
times out.  If no transition can leave that state, the relay is
wedged --- every tool call blocks on \texttt{nats: timeout} with
no recovery (biff-tww).  This specification adds the half-open
state explicitly and proves a liveness property: the connection
can never be stuck there.  Two independent transitions escape the
wedge.  The keepalive detector ($WedgeDetected$) waits for
nats-py's ping loop to declare the socket dead and hands it to the
background reconnect; the proactive detector ($ForceReconnect$,
biff-3hp) force-closes the socket after three consecutive request
timeouts and drops the client, so the next lazy connect dials a
fresh one.  The proof is by deadlock analysis --- in the absence
of \emph{any} wedge-detection transition, probcli finds half-open
to be a reachable terminal state, and that terminal state
\emph{is} biff-tww.  With either recovery transition present,
half-open is deadlock-free.

A third invariant, added here, concerns a resource the earlier model
did not track at all: the always-on core-NATS talk subscription
(biff-9la).  The relay opens one long-lived SUB on the talk inbox and
holds it for the life of the process.  A core-NATS subscription is
bound to the particular client that created it and dies with that
client, so the proactive force-reconnect ($ForceReconnect$, biff-3hp),
which discards the client wholesale, silently orphans it: presence and
wall recover on the fresh connection, but talk stays dead for the rest
of the session because nothing re-opens the SUB.  This is the exact
converse of nats-py's in-place reconnect, which keeps the client and
replays its subscriptions, so the SUB must be \emph{preserved} there
and \emph{re-established} after a force-reconnect --- the same
distinction the handle-rebuild machinery already draws, but for a
resource the earlier model omitted.  The specification adds the
subscription's generation binding as state and proves a liveness
property: a live connection whose talk SUB is stranded on a superseded
generation always eventually re-binds it.  The defect is reproduced as
a ProB counterexample against the pre-fix ``re-subscribe only when the
handle is \texttt{None}'' guard, and the generation-tracked guard makes
the stranded state recover.

A fourth invariant, added here, closes the converse of the wedge race:
not a failure to \emph{escape} the wedge, but an unnecessary
\emph{teardown} of a connection that has already left it.  The proactive
force-reconnect ($ForceReconnect$, biff-3hp) is a two-phase act --- a
gate that decides, after three request timeouts, that the socket is
wedged, and a scheduled coroutine that later takes the lock and
force-closes it.  Between the two, nats-py's keepalive can complete a
\emph{transparent} reconnect on the \emph{same} client object: the
socket the gate condemned is healthy again, with no new dial and so no
change to \texttt{\_generation}.  The scheduled teardown, re-checking
under the lock, sees \texttt{is\_connected} true --- exactly as it was
during the wedge --- and, unable to tell a healed connection from a
still-wedged one, force-closes a freshly-healthy client (biff-xko).  The
cost is one unnecessary reconnect-and-reprovision ($\sim$20\,ms, no data
loss), but it recurs, and it is the same class of half-open
indistinguishability the earlier fixes met from the other side.
\texttt{is\_connected} cannot separate the two states because it is true
in both; the model separates them with $reconnectEpoch$, a per-client
counter of transparent reconnects that advances where
\texttt{\_generation} does not.  This specification adds that counter as
state and proves a safety property: a scheduled force-reconnect whose
gate observed an earlier epoch is a no-op against a client that has since
reconnected --- the live connection is invariant under the stale
teardown.  The defect is reproduced as a ProB counterexample against the
pre-fix \texttt{is\_connected}-only re-check (the frozen variant
\texttt{nats-relay-xko-isconnected.tex}), and the epoch-guarded re-check
makes the spurious teardown unreachable while leaving the force-reconnect
of a genuinely wedged client intact.

A fifth refinement, added here, corrects an over-idealisation in the
fourth.  The epoch guard just described assumes nats-py's transparent
reconnect is atomic in $(connState, reconnectEpoch)$ --- that the socket
comes up and the epoch advances in one step, so an under-lock re-check
that reads \texttt{is\_connected} true always also reads the advanced
epoch.  It is not atomic.  In \texttt{nats/aio/client.py} the reconnect
coroutine sets the status to \texttt{CONNECTED}, then \texttt{await
self.flush()}, and only afterwards fires \texttt{reconnected\_cb} ---
biff's \texttt{\_on\_reconnect}, which increments the epoch.  There is a
reachable interleaving in which the socket is up but the epoch has not yet
been bumped, and a scheduled force-teardown, re-checking under the lock,
sees \texttt{is\_connected} true and \texttt{\_reconnect\_epoch ==
gate\_epoch} --- both conjuncts the fix relied on to skip --- and so does
\emph{not} skip: it tears a freshly-healed client down.  The deferred
\texttt{\_on\_reconnect} then lands on the reset epoch, momentarily
leaving $reconnectEpoch = 1$ while the connection is $disconnected$ and
the client is $None$, which the fourth model's state invariant $connState
\in \{ disconnected, closed \} \implies reconnectEpoch = 0$ wrongly
declared impossible.  This specification adds the flush window as an
explicit state ($epochPending$), splits the atomic reconnect into
$ReconnectSocketUp$ and $ReconnectEpochObserved$, models the flush-window
teardown as $FlushWindowForceTeardown$, and reclassifies the epoch-reset
invariant from a global always-property to an \emph{eventually}-true
recovery property.  It then proves what matters: the flush-window teardown
of a healthy client is bounded and self-healing --- every trace through it
returns to a healthy $connected$ state via a fresh dial --- the biff-tww
deadlock remains unreachable, and no new bad state is introduced.  The
epoch guard is not weakened; it is shown to eliminate exactly the
teardowns it \emph{can} observe, with the residual, unobservable window
proved harmless rather than hidden.

%%----------------------------------------------------------------------
\section{Basic Types}
%%----------------------------------------------------------------------

\begin{zed}
  [USER, TTY, REPO, KVKEY, KVVALUE, SESSIONKEY]
\end{zed}

$USER$ identifies a human user.  $TTY$ identifies a terminal
session (random 8-char hex).  $REPO$ scopes all state to a
repository.  $KVKEY$ and $KVVALUE$ are opaque KV store keys and
values.  $SESSIONKEY$ is the composite $\{user\}:\{tty\}$ identifier.

%%----------------------------------------------------------------------
\section{Free Types}
%%----------------------------------------------------------------------

\begin{zed}
  ZBOOL ::= ztrue | zfalse
\end{zed}

\begin{zed}
  ConnState ::= disconnected | reconnecting | connected | halfOpen | closed
\end{zed}

The connection has five states.  $disconnected$ is a deliberate,
reversible release of the TCP connection (the relay's own
\texttt{disconnect()}); a subsequent call reconnects.
$reconnecting$ is distinct: nats-py's background reconnect loop is
in progress after a dropped socket --- \texttt{is\_closed} is
false, \texttt{is\_connected} is false, and cached handles have
already been invalidated by \texttt{\_on\_disconnect}.
$connected$ is an active TCP connection whose JetStream and KV
handles are provisioned.  $halfOpen$ is the wedge: the socket is
up (\texttt{is\_closed} false) but every JetStream or KV request
times out; the cached handles are stale and must not be used.
$closed$ is permanent shutdown --- no reconnection is possible.

\begin{zed}
  ReqOutcome ::= reqOk | reqTimeout
\end{zed}

The outcome of a JetStream or KV request issued on a $connected$
connection.  $reqOk$ is a normal response.  $reqTimeout$ is a
\texttt{nats.errors.TimeoutError}: the request was sent but no
reply arrived within the deadline, which is the observable
signature of a half-open connection.

\begin{zed}
  WatcherPhase ::= idle | snapshotting | streaming | stopped
\end{zed}

The KV watcher progresses through phases: $idle$ (not yet started),
$snapshotting$ (receiving the initial KV snapshot), $streaming$
(receiving live updates post-snapshot), and $stopped$ (shutdown
complete).

\begin{zed}
  KVOp ::= kvPut | kvDelete | kvPurge
\end{zed}

Operations observed on KV entries.  $kvPut$ stores or updates a
value.  $kvDelete$ and $kvPurge$ remove entries (TTL expiry
produces $kvDelete$).

\begin{zed}
  WatcherEvent ::=
  weEntry | weSnapshotDone | weTimeout | weError | weShutdown
\end{zed}

Events that the KV watcher loop must handle.  $weEntry$ is a KV
change.  $weSnapshotDone$ is the $None$ marker from
\texttt{watcher.updates()}.  $weTimeout$ is a
\texttt{TimeoutError} from idle periods.  $weError$ is an
unrecoverable NATS error.  $weShutdown$ is the shutdown signal.

%%----------------------------------------------------------------------
\section{Global Constants}
%%----------------------------------------------------------------------

\begin{axdef}
  kvTtlSeconds : \nat \\
  watcherTimeoutSeconds : \nat \\
  maxSessions : \nat \\
  restartDelaySeconds : \nat \\
  maxGeneration : \nat \\
  maxEpoch : \nat
  \where
  kvTtlSeconds = 259200 \\
  watcherTimeoutSeconds = 5 \\
  maxSessions = 1000 \\
  restartDelaySeconds = 2 \\
  maxGeneration = 3 \\
  maxEpoch = 2
\end{axdef}

$kvTtlSeconds$ is the NATS KV TTL (3 days).
$watcherTimeoutSeconds$ is the \texttt{updates()} timeout that
keeps the shutdown check responsive.  $maxSessions$ bounds the
model for animation.  $maxGeneration$ bounds the per-dial
generation counter (below) so the reachable state space stays
finite; three is enough to exhibit a superseded generation
alongside the live one.  $maxEpoch$ bounds the per-client
\emph{in-place reconnect} counter $reconnectEpoch$ (below), which
counts nats-py's transparent same-object reconnects rather than
fresh dials; two is enough to exhibit a reconnect that completes
\emph{after} a force-reconnect gate has observed the pre-reconnect
value (gate epoch $0$, $reconnectEpoch = 1$), which is the whole
of the biff-xko race.

%%----------------------------------------------------------------------
\section{Connection State}
%%----------------------------------------------------------------------

\begin{schema}{Connection}
  connState : ConnState \\
  kvProvisioned : ZBOOL \\
  streamsProvisioned : ZBOOL \\
  namesProvisioned : ZBOOL \\
  wtmpAvailable : ZBOOL \\
  generation : \nat \\
  talkSubGen : \nat \\
  reconnectEpoch : \nat \\
  spuriousTeardown : ZBOOL \\
  epochPending : ZBOOL
  \where
  connState \neq connected \implies
  kvProvisioned = zfalse \land
  streamsProvisioned = zfalse \land
  namesProvisioned = zfalse \\
  connState = connected \implies
  kvProvisioned = ztrue \land
  streamsProvisioned = ztrue \land
  namesProvisioned = ztrue \\
  connState \in \{ connected, reconnecting, halfOpen \} \implies
  generation \geq 1 \\
  generation \leq maxGeneration \\
  talkSubGen \leq generation \\
  reconnectEpoch \leq maxEpoch \\
  epochPending = ztrue \implies
  connState \in \{ connected, disconnected \}
\end{schema}

This is the generalised handle-validity invariant: valid
JetStream, sessions-KV, and names-KV handles exist \emph{if and
only if} the connection is $connected$.  In particular they do
not exist in $halfOpen$ --- the stale handles cached during the
wedge are, in the model, no handles at all.  This generalises the
earlier ``$disconnected$ or $closed \implies$ no handles'' (DES-029)
to cover $reconnecting$ and $halfOpen$ as well.  It is the
formal statement that a half-open connection must never be treated
as usable: the biff-tww crash-loop arose precisely because the code
kept and re-used handles that this invariant forbids.
$wtmpAvailable$ degrades independently --- wtmp provisioning
failure is non-fatal --- and is cleared whenever the handles are.

$generation$ is the formal counterpart of the relay's per-dial
token \texttt{\_generation}.  Each fresh \texttt{nats.connect}
(the slow path of \texttt{\_ensure\_connected}, taken when the
client was dropped) bumps it; nats-py's own in-place reconnect,
which reuses the same client, does not.  It exists so the model
can distinguish a request settling on the \emph{current} client
from one settling on a \emph{superseded} client that a rebuild has
already replaced --- the race $ForceReconnect$'s identity guard
must survive.  A settle carries the generation it was issued on;
comparing that to $generation$ is the model's \texttt{self.\_nc is
owner} check.  It is bounded by $maxGeneration$ so the state space
stays finite.

The added invariant $connState \in \{ connected, reconnecting,
halfOpen \} \implies generation \geq 1$ (equivalently $generation =
0 \implies connState \in \{ disconnected, closed \}$) makes
machine-checked the fact that reaching a live or recovering state
requires at least one dial.  It holds by induction --- $generation$
starts at $0$ in $disconnected$ and the only path to $connected$ is
$EnsureConnected$ from $disconnected$, which bumps it to at least
$1$ --- but stating it as an invariant makes machine-checked the one
fact $StaleRequestTimeout$ needs: a live state has $generation \geq 1$,
so a strictly-smaller $owner?$ value exists.  It does \emph{not} prove
that a genuinely-superseded live generation was ever issued --- $owner?$
is a free $\nat$ the model never ties to an actually-dialled generation,
so the precondition $owner? < generation$ can name a value that was never
live.  This is sound only because $StaleRequestTimeout$ is a $\Xi$ no-op
on connection state: a stale settle changes nothing, so a spurious
$owner?$ cannot corrupt anything.  The invariant also guards against a
future edit that reaches $connected$ without a dial.

$talkSubGen$ is the generation to which the always-on core-NATS talk
subscription is currently bound.  The relay opens one long-lived SUB on
the talk inbox (biff-9la) and holds it for the life of the process; a
core-NATS subscription is bound to the \emph{particular} client that
created it, and so it dies with that client.  A value of $0$ means no
SUB has been established on any dial --- the initial state, and the
window between a fresh connect and the poller's first subscribe.  A
value $g \geq 1$ means the live SUB was subscribed on the client of
generation $g$.  The subscription is \emph{current} exactly when
$talkSubGen = generation$ and \emph{stranded} --- bound to a client the
rebuild has already replaced, so no talk message will ever arrive ---
when $talkSubGen < generation$.  The invariant $talkSubGen \leq
generation$ records the one thing that can never happen: a SUB cannot be
bound to a client that was never dialled.  It is the exact analogue of
the handle-provisioned flags, but where those are cleared and
re-established on \emph{every} reconnect, the talk SUB is preserved
across nats-py's in-place reconnect (same client, $generation$
unchanged, SUBs replayed) and re-established only when a fresh dial
advances $generation$ past $talkSubGen$.  This is the distinction the
biff-9la bug missed: a force-reconnect (biff-3hp) replaces the client
and orphans the SUB, and unless the poller re-subscribes on the new
generation, talk dies silently for the rest of the session.  The
stranded state $connState = connected \land talkSubGen < generation$ is
reachable and legal --- it is the transient window before the poller
ticks --- so the property that forbids the bug is not a state invariant
but the liveness obligation that the stranding is always eventually
lifted (\S\ref{sec:modelcheck}).

$reconnectEpoch$ is the resource $generation$ deliberately does not
count: nats-py's \emph{transparent, same-object} reconnect.  When the
keepalive loop declares a socket dead and its background reconnect
succeeds, nats-py fires \texttt{\_on\_reconnect} on the \emph{same}
\texttt{Client} object --- no new \texttt{nats.connect}, so $generation$
is unchanged --- and replays the client's subscriptions.  Each such
completion advances $reconnectEpoch$ (saturating at $maxEpoch$).  It is
reset to $0$ whenever the client is discarded: on a fresh dial, on
$Disconnect$, on $Close$, and on the force-reconnect teardown.  The pair
$(generation, reconnectEpoch)$ is thus the model's complete identity of
the live transport: $generation$ changes exactly on a new dial,
$reconnectEpoch$ exactly on an in-place reconnect, and the two never
change together.  This is the distinction \texttt{is\_connected} cannot
draw and the whole of what the biff-xko fix turns on.

An earlier version of this specification pinned the reset with the state
invariant $connState \in \{ disconnected, closed \} \implies
reconnectEpoch = 0$: a counter of reconnects on the current client is
meaningless once there is no current client.  That invariant is
\emph{false} of the code, and the correction is the subject of
\S\ref{sec:flushwindow}.  nats-py's in-place reconnect is not atomic in
$(connState, reconnectEpoch)$.  It sets the socket status to $connected$,
then performs a network flush, and only \emph{afterwards} fires
\texttt{\_on\_reconnect} --- the callback in which the relay executes
\texttt{self.\_reconnect\_epoch += 1}.  A scheduled force-teardown can
run in that flush window, discard the client (resetting the epoch to
$0$), and then have the deferred \texttt{\_on\_reconnect} land on the
already-reset counter, leaving $reconnectEpoch = 1$ while $connState =
disconnected$ and the client is $None$: exactly the state the old
invariant forbids.  The reset is therefore not a global always-property
but an \emph{eventually}-true recovery property --- the next dial resets
the epoch --- proved as a $CTL$ formula in \S\ref{sec:modelcheck} rather
than asserted over every state.

$epochPending$ is the state that makes the flush window visible.  It is
$ztrue$ from the moment nats-py's reconnect brings the socket up
($ReconnectSocketUp$: $connState = connected$, but $reconnectEpoch$ not
yet advanced) until the deferred \texttt{\_on\_reconnect} lands and bumps
the epoch ($ReconnectEpochObserved$).  It is the model's record of a
transparent reconnect that has \emph{completed at the socket} but has
\emph{not yet been observed} through the epoch --- the one thing the
under-lock re-check cannot know, because \texttt{is\_connected} is $ztrue$
throughout it and $reconnectEpoch$ has not moved.  While it is set, the
epoch guard $gateEpoch? < reconnectEpoch$ is \emph{false} (the epoch still
equals the gate's), so the re-check does not skip: it tears the
freshly-healed client down.  The invariant $epochPending = ztrue \implies
connState \in \{ connected, disconnected \}$ records that a pending epoch
observation exists only where a client (or its just-discarded remains)
does: in the flush window itself ($connected$) or in the brief interval
after a flush-window teardown but before the deferred bump lands and the
next dial supersedes it ($disconnected$).

The force-reconnect decision itself --- the \emph{gate} ---
is not carried in the state.  The relay's \texttt{\_tracked} wrapper,
after three consecutive request timeouts on a still-\texttt{is\_connected}
socket, decides to force-reconnect and schedules
\texttt{\_force\_reconnect}; only later does that coroutine take the
connection lock and re-check before tearing the client down.  Between
the decision and the teardown an arbitrary interleaving is possible ---
and that window is the biff-xko race.  Following the treatment of the
DES-042 identity guard already in this model, the value the gate
\emph{observed} is modelled not as a stored flag but as the input
$gateEpoch?$ of the re-check operation $StaleForceTeardown$ below,
exactly as $StaleRequestTimeout$ carries the superseded generation as
the input $owner?$ rather than as connection state.  The gate is a
latency parameter of the detector, recorded in prose as
\texttt{ping\_interval} and the three-timeout threshold are; what the
model must make checkable is only the \emph{comparison} the under-lock
re-check performs, and that comparison lives in the guard of the
re-check, against the input the gate supplied.

$spuriousTeardown$ is not relay state but a \emph{ghost observer}.  It
starts $zfalse$, every genuine transport operation preserves it, and it
is flipped to $ztrue$ only by a re-check that tears down a client which
has \emph{detectably} reconnected since the gate --- exactly the
biff-xko event: an unnecessary teardown of a freshly-healthy connection.
Deliberately it is \emph{not} pinned by the state invariant; it is a
free boolean the model-checker carries so that the safety property can
be stated over reachable states as the temporal formula
$AG\,(spuriousTeardown = zfalse)$ (\S\ref{sec:modelcheck}), rather than
as a latch invariant.  A latch invariant would be unsound against the
constraint-based deadlock check, which explores invariant-violating
candidate states and would read a pinned-but-set observer as a dead
sink; a $CTL$ property over reachable states has no such hazard.  In the
shipped model no operation can flip the observer --- the re-check
($StaleForceTeardown$) is a $\Xi$ no-op --- so $AG\,(spuriousTeardown =
zfalse)$ is $TRUE$.  In the frozen pre-fix variant
\texttt{nats-relay-xko-isconnected.tex}, whose re-check reads only
\texttt{is\_connected} and tears the healed client down, the observer is
flipped and probcli returns the flip as the counterexample.  The
observer is not idle scaffolding: it is the only way to distinguish a
legitimate force-reconnect of a genuinely wedged client, which also
lands in $disconnected$, from the spurious teardown of a healed one,
since the two share an after-state and differ only in the transition
that reached it.

\begin{schema}{InitConnection}
  Connection'
  \where
  connState' = disconnected \\
  kvProvisioned' = zfalse \\
  streamsProvisioned' = zfalse \\
  namesProvisioned' = zfalse \\
  wtmpAvailable' = zfalse \\
  generation' = 0 \\
  talkSubGen' = 0 \\
  reconnectEpoch' = 0 \\
  spuriousTeardown' = zfalse \\
  epochPending' = zfalse
\end{schema}

%%----------------------------------------------------------------------
\section{Connection Operations}
%%----------------------------------------------------------------------

\subsection{Ensure Connected (Lazy Connect)}

\begin{schema}{EnsureConnected}
  \Delta Connection
  \where
  connState = disconnected \lor connState = connected \\
  connState' = connected \\
  kvProvisioned' = ztrue \\
  streamsProvisioned' = ztrue \\
  namesProvisioned' = ztrue \\
  wtmpAvailable' = wtmpAvailable \\
  ( connState = disconnected \land
    ( ( generation < maxGeneration \land
        generation' = generation + 1 ) \lor
      ( generation = maxGeneration \land
        generation' = generation ) ) \land
    reconnectEpoch' = 0 \land epochPending' = zfalse ) \lor \\
  ( connState = connected \land generation' = generation \land
    reconnectEpoch' = reconnectEpoch \land epochPending' = epochPending ) \\
  talkSubGen' = talkSubGen \\
  spuriousTeardown' = spuriousTeardown
\end{schema}

The relay's lazy connect (\texttt{\_ensure\_connected}).  It is
enabled only from $disconnected$ (a fresh connect) or $connected$
(idempotent no-op).  The fresh connect from $disconnected$ dials a
new client and so bumps $generation$ (saturating at
$maxGeneration$ to keep the model finite); the idempotent no-op
from $connected$ reuses the live client and leaves $generation$
unchanged.  It is \emph{not} enabled from $halfOpen$:
this is the formal counterpart of the DES-029 bug.  The fast path
sees \texttt{is\_closed} false and returns the stale cached
handles without reconnecting, so $EnsureConnected$ makes no
progress out of the wedge and must not be modelled as if it did.
Nor is it enabled from $reconnecting$ (nats-py owns that window;
provisioning would block until reconnect completes, bounded by
\texttt{\_CONNECT\_PROVISION\_TIMEOUT}, biff-wr3) or from $closed$
(no reconnection after close).  The fresh dial resets $reconnectEpoch$
to $0$ (the relay's \texttt{self.\_reconnect\_epoch = 0} in
\texttt{\_open\_connection}): the new client has undergone no in-place
reconnects, and this is the transition that heals the transient
$reconnectEpoch > 0$ state a flush-window teardown can leave behind
(\S\ref{sec:flushwindow}); it also clears $epochPending$, since a fresh
dial bumps $generation$ and so supersedes any deferred
\texttt{\_on\_reconnect} bound to the discarded client.  The idempotent
no-op from $connected$ preserves both.  $wtmpAvailable$ is preserved ---
it only changes during provisioning.  $talkSubGen$ is likewise
preserved: dialling a fresh client bumps $generation$ but does
\emph{not} re-open the talk SUB.  This is deliberate, and it is where
the stranding becomes observable.  When the slow path fires from
$disconnected$, $generation$ advances while $talkSubGen$ stays on the
old value, so the post-connect state is $connState = connected$ with
$talkSubGen < generation$ --- a live connection whose talk SUB is bound
to the client that was just discarded.  Re-establishing the SUB is not
the connect's job but the poller's, modelled by $Subscribe$ below; the
model thus separates ``the client came back'' from ``the talk SUB was
rebound to it,'' which is exactly the separation the biff-9la bug
elided.

\subsection{Disconnect (Reversible)}

\begin{schema}{Disconnect}
  \Delta Connection
  \where
  connState = connected \\
  connState' = disconnected \\
  kvProvisioned' = zfalse \\
  streamsProvisioned' = zfalse \\
  namesProvisioned' = zfalse \\
  wtmpAvailable' = zfalse \\
  generation' = generation \\
  talkSubGen' = talkSubGen \\
  reconnectEpoch' = 0 \\
  spuriousTeardown' = spuriousTeardown \\
  epochPending' = zfalse
\end{schema}

The relay's explicit, reversible \texttt{disconnect()}: it acts
on a healthy $connected$ connection and releases it.  The next
relay call reconnects via $EnsureConnected$.  It is deliberately
\emph{not} the escape from $halfOpen$ --- the running server does
not autonomously call \texttt{disconnect()} while wedged.  The
wedge is left only by autonomous detection: the keepalive path
($WedgeDetected$) or the proactive force-close ($ForceReconnect$),
never by an ordinary \texttt{disconnect()}.  Note, however, that
$disconnected$ is the state $ForceReconnect$ moves \emph{to}: a
force-closed socket with its client dropped is, like a deliberate
\texttt{disconnect()}, a connection whose next call must dial
afresh.

\subsection{Close (Permanent)}

\begin{schema}{Close}
  \Delta Connection
  \where
  connState \neq halfOpen \\
  connState' = closed \\
  kvProvisioned' = zfalse \\
  streamsProvisioned' = zfalse \\
  namesProvisioned' = zfalse \\
  wtmpAvailable' = zfalse \\
  generation' = generation \\
  talkSubGen' = talkSubGen \\
  reconnectEpoch' = 0 \\
  spuriousTeardown' = spuriousTeardown \\
  epochPending' = zfalse
\end{schema}

Permanent shutdown, reachable from any \emph{definite} transport
state.  The precondition $connState \neq halfOpen$ records a
modelling decision that is load-bearing for the liveness proof:
an orderly close cannot be the escape from the wedge.  In the
biff-tww crash-loop the server never reaches an orderly
\texttt{close()} --- it is blocked awaiting a request that will
never return --- so $halfOpen$ must be escaped by autonomous
wedge detection, not by $Close$.  From $closed$ the operation is
an idempotent self-loop, which keeps $closed$ deadlock-free.

\subsection{Reset Infrastructure}

\begin{schema}{ResetInfrastructure}
  \Delta Connection
  \where
  connState \neq halfOpen \\
  connState' = connState \\
  kvProvisioned' = zfalse \\
  streamsProvisioned' = zfalse \\
  namesProvisioned' = zfalse \\
  wtmpAvailable' = zfalse \\
  generation' = generation \\
  talkSubGen' = talkSubGen \\
  reconnectEpoch' = reconnectEpoch \\
  spuriousTeardown' = spuriousTeardown \\
  epochPending' = epochPending
\end{schema}

Clears cached handles without changing the transport state
(\texttt{reset\_infrastructure()}).  By the handle-validity
invariant, $connState' = connState$ together with cleared handles
forces $connState \neq connected$; the operation is therefore a
self-loop at $disconnected$, $reconnecting$, and $closed$, where
the handles are already absent.  It is excluded from $halfOpen$
for the same reason as $Close$: clearing a cached handle does not
lift the wedge (the socket is still up and still unresponsive),
so it must not be modelled as an escape.

\subsection{Request (JetStream / KV Round-Trip)}

\begin{schema}{Request}
  \Delta Connection \\
  result! : ReqOutcome
  \where
  connState = connected \\
  generation' = generation \\
  talkSubGen' = talkSubGen \\
  reconnectEpoch' = reconnectEpoch \\
  spuriousTeardown' = spuriousTeardown \\
  ( result! = reqOk \land
    connState' = connected \land
    kvProvisioned' = ztrue \land
    streamsProvisioned' = ztrue \land
    namesProvisioned' = ztrue \land
    wtmpAvailable' = wtmpAvailable \land
    epochPending' = epochPending ) \\
  \lor \\
  ( result! = reqTimeout \land
    connState' = halfOpen \land
    kvProvisioned' = zfalse \land
    streamsProvisioned' = zfalse \land
    namesProvisioned' = zfalse \land
    wtmpAvailable' = zfalse \land
    epochPending' = zfalse )
\end{schema}

Every JetStream or KV round-trip issued \emph{and settled} on the
current client --- \texttt{stream\_info} in
\texttt{get\_unread\_summary}, \texttt{kv.get} in
\texttt{heartbeat}, and the like.  From $connected$ it has two
outcomes.  $reqOk$ leaves the connection $connected$ with its
handles intact.  $reqTimeout$ is the half-open signature: the
request timed out, the connection moves to $halfOpen$, and the
handles are invalidated to satisfy the handle-validity invariant.
This is the operation that makes $halfOpen$ reachable, and the
crash-loop of biff-tww is exactly $reqTimeout$ fired repeatedly
with no way out.

Because $Request$ both issues and settles on the current client,
its $reqTimeout$ is the \emph{current-generation} timeout ---
$\texttt{self.\_nc is owner}$ holds --- so it legitimately drives
the wedge.  The distinct case where the client was replaced
between issue and settle is $StaleRequestTimeout$ below, which does
\emph{not} drive the wedge.  $generation$ is preserved either way:
a round-trip never dials a new client.

$epochPending$ is preserved on $reqOk$ (a healthy round-trip, whether or
not the connection is mid-flush-window) and cleared on $reqTimeout$.  The
clear is what keeps the flush window from leaking into the wedge: a fresh
timeout drives $connected \to halfOpen$, a new wedge episode whose
recovery is a new reconnect cycle, so any epoch observation still pending
from the \emph{previous} reconnect is scoped out here.  Because
$reqTimeout$ is the only edge into $halfOpen$, this clear is exactly what
makes $halfOpen$ --- and hence $reconnecting$, reached only from
$halfOpen$ by $WedgeDetected$ --- always carry $epochPending = zfalse$,
which is the precondition $ReconnectSocketUp$ relies on to be enabled.

\subsection{Wedge Detected (Recovery)}

\begin{schema}{WedgeDetected}
  \Delta Connection
  \where
  connState = halfOpen \\
  connState' = reconnecting \\
  kvProvisioned' = zfalse \\
  streamsProvisioned' = zfalse \\
  namesProvisioned' = zfalse \\
  wtmpAvailable' = zfalse \\
  generation' = generation \\
  talkSubGen' = talkSubGen \\
  reconnectEpoch' = reconnectEpoch \\
  spuriousTeardown' = spuriousTeardown \\
  epochPending' = zfalse
\end{schema}

The keepalive-driven recovery transition.
Keepalive PINGs (\texttt{ping\_interval=20}, \texttt{max\_outstanding\_pings=3})
detect that the server has stopped answering within
$\sim$60\,s and fire nats-py's \texttt{\_on\_disconnect}, which
invalidates the cached handles and hands the socket to nats-py's
infinite reconnect loop.  Before the proactive detector this was
the \emph{only} operation enabled in $halfOpen$; removing it (the
pre-fix code, whose \texttt{ping\_interval} default of 240\,s never
fired in practice) made $halfOpen$ a terminal state, and probcli
surfaced that terminal state --- biff-tww --- as a reachable
deadlock (the frozen \texttt{nats-relay-passA-nowedge} artifacts).
$ForceReconnect$ (below) now adds a second, faster escape from the
same state; with either operation present, $halfOpen$ is
deadlock-free.

\subsection{Reconnect Socket Up (nats-py Reconnect Reaches CONNECTED)}
\label{sec:flushwindow}

nats-py's transparent reconnect is not a single step.  In
\texttt{nats/aio/client.py} the reconnect coroutine sets
\texttt{self.\_status = CONNECTED} (the socket is up,
\texttt{is\_connected} becomes $ztrue$), then \texttt{await
self.flush()} --- a network round-trip --- and only afterwards fires
\texttt{reconnected\_cb}, the relay's \texttt{\_on\_reconnect}, in which
\texttt{self.\_reconnect\_epoch += 1} runs.  The socket coming up and the
epoch advancing are therefore \emph{two} steps with an \texttt{await}
between them, and an arbitrary interleaving is possible in the flush
window.  The earlier model collapsed both into one atomic $Reconnect$;
that atomicity is exactly the divergence this section corrects.  The
completion is split into $ReconnectSocketUp$ (the socket reaches
$connected$, epoch unchanged, observation pending) and
$ReconnectEpochObserved$ (the deferred \texttt{\_on\_reconnect} lands and
advances the epoch).  A force-teardown may interleave between them.

\begin{schema}{ReconnectSocketUp}
  \Delta Connection
  \where
  connState = reconnecting \\
  epochPending = zfalse \\
  connState' = connected \\
  kvProvisioned' = ztrue \\
  streamsProvisioned' = ztrue \\
  namesProvisioned' = ztrue \\
  wtmpAvailable' = wtmpAvailable \\
  generation' = generation \\
  talkSubGen' = talkSubGen \\
  reconnectEpoch' = reconnectEpoch \\
  spuriousTeardown' = spuriousTeardown \\
  epochPending' = ztrue
\end{schema}

nats-py brings the socket back up on the \emph{same} client and sets its
status to CONNECTED; the next relay call re-provisions the handles.
$generation$ is unchanged --- no new \texttt{nats.connect} --- so the
per-dial token does not advance.  Crucially $reconnectEpoch$ is
\emph{also} unchanged: the epoch bump lives in
\texttt{\_on\_reconnect}, which has not yet fired.  The step records that
fact by setting $epochPending' = ztrue$.  This closes the socket half of
the recovery path $connected \to halfOpen \to reconnecting \to
connected$, so no reachable state is stuck in $halfOpen$; the epoch half
follows in $ReconnectEpochObserved$.  The precondition $epochPending =
zfalse$ is discharged for free: $reconnecting$ is entered only by
$WedgeDetected$ from $halfOpen$, and $halfOpen$ always carries
$epochPending = zfalse$ (the $Request(reqTimeout)$ edge that is its only
entry clears it), so the step is always enabled in $reconnecting$ and
$reconnecting$ never deadlocks.

$talkSubGen$ is preserved: the SUB was created on this very client, and
nats-py replays it on reconnect.  Because $generation$ is unchanged, a
SUB that was current before the wedge ($talkSubGen = generation$) is
current after it, with no re-subscribe required.  This is the
load-bearing contrast with $ForceReconnect$: the keepalive path keeps the
client, so the talk SUB survives; the force-close path discards it, so
the talk SUB must be rebuilt.

\subsection{Reconnect Epoch Observed (Deferred \texttt{\_on\_reconnect}
Lands)}

\begin{schema}{ReconnectEpochObserved}
  \Delta Connection
  \where
  epochPending = ztrue \\
  connState \in \{ connected, disconnected \} \\
  connState' = connState \\
  kvProvisioned' = kvProvisioned \\
  streamsProvisioned' = streamsProvisioned \\
  namesProvisioned' = namesProvisioned \\
  wtmpAvailable' = wtmpAvailable \\
  generation' = generation \\
  talkSubGen' = talkSubGen \\
  ( ( reconnectEpoch < maxEpoch \land
      reconnectEpoch' = reconnectEpoch + 1 ) \lor
    ( reconnectEpoch = maxEpoch \land
      reconnectEpoch' = reconnectEpoch ) ) \\
  spuriousTeardown' = spuriousTeardown \\
  epochPending' = zfalse
\end{schema}

The deferred \texttt{\_on\_reconnect} fires and runs
\texttt{self.\_reconnect\_epoch += 1} (saturating at $maxEpoch$),
clearing the pending observation.  This is the single most important
transition for biff-xko: it is the model's record that a transparent
same-object reconnect has been \emph{observed}.  Because $generation$ is
fixed while $reconnectEpoch$ ticks, a force-reconnect whose gate fired
before it --- observing the pre-reconnect epoch $gateEpoch?$ --- can now
determine, under its lock, that $gateEpoch? < reconnectEpoch$: the wedge
it was dispatched to tear down has already healed.  This is the
observation \texttt{is\_connected} cannot make, because
\texttt{is\_connected} is $ztrue$ across the whole window; the epoch
separates healed from wedged, \texttt{is\_connected} conflates them.

The operation is enabled from $connected$ \emph{and} from $disconnected$,
and the second case is the crux of the divergence.  The normal
completion is from $connected$: the flush window resolves with the client
still live, the epoch advances to $1$, and the connection is fully
healed.  But if a $FlushWindowForceTeardown$ (below) has already fired
--- discarding the client, dropping to $disconnected$, and resetting the
epoch to $0$ --- the deferred \texttt{\_on\_reconnect}, still bound to the
un-superseded $generation$, lands \emph{after} the teardown and bumps the
reset counter to $reconnectEpoch = 1$ while $connState = disconnected$.
That is the state the old always-invariant $connState \in \{
disconnected, closed \} \implies reconnectEpoch = 0$ declared impossible,
and it is genuinely reachable in the code (\texttt{\_force\_reconnect}
resets the epoch to $0$; the late \texttt{\_on\_reconnect} increments it).
The model states it honestly; the recovery property in
\S\ref{sec:modelcheck} proves it transient --- the next $EnsureConnected$
dial resets the epoch and re-supersedes the callback.

\subsection{Force Reconnect (Proactive Wedge Teardown)}

\begin{schema}{ForceReconnect}
  \Delta Connection
  \where
  connState = halfOpen \\
  connState' = disconnected \\
  kvProvisioned' = zfalse \\
  streamsProvisioned' = zfalse \\
  namesProvisioned' = zfalse \\
  wtmpAvailable' = zfalse \\
  generation' = generation \\
  talkSubGen' = talkSubGen \\
  reconnectEpoch' = 0 \\
  spuriousTeardown' = spuriousTeardown \\
  epochPending' = zfalse
\end{schema}

The proactive escape from the wedge (\texttt{\_force\_reconnect},
biff-3hp), and the outcome of an under-lock re-check that finds the
client \emph{still wedged}.  Where $WedgeDetected$ waits for nats-py's
keepalive to declare the socket dead --- a $\sim$60--80\,s floor ---
this operation acts on the relay's own evidence: after
\texttt{\_WEDGE\_FORCE\_RECONNECT\_THRESHOLD} $= 3$ consecutive runtime
request timeouts on a still-\texttt{is\_connected} socket,
\texttt{\_tracked} schedules \texttt{\_force\_reconnect}, which takes the
lock, confirms the client is the same wedged one (no reconnect has
intervened --- the condition made precise in $StaleForceTeardown$
below), calls \texttt{safe\_close(nc)}, and drops the client.  Each
timed-out request blocks $\sim$5\,s, so the bound is $3 \times
5\,\mathrm{s} \approx 15\,$s --- four to five times faster than the
keepalive path.  It is modelled atomically because the still-wedged case
has no interesting intermediate state: the socket the gate condemned is
the socket under the lock.  Enabled unconditionally from $halfOpen$, it
is also the second escape from the wedge (with $WedgeDetected$), so
$halfOpen$ is never terminal.

The target is $disconnected$, not $reconnecting$ --- the load-bearing
distinction between the two escapes.  $WedgeDetected$ leaves the
\emph{same} client in nats-py's hands and recovers through
$reconnecting$ by $Reconnect$; $ForceReconnect$ discards the client
entirely, so there is no socket for nats-py to reconnect and recovery is
the existing lazy connect from $disconnected$.  The full proactive path
is $connected \to halfOpen \to disconnected \to connected$, closed by
operations the model already contains.  $reconnectEpoch$ is reset to $0$
--- the discarded client's in-place reconnect count dies with it --- and
the three-consecutive-timeout threshold is not modelled as a counter,
for the same symmetry reason the keepalive's three-missed-PING criterion
is abstracted into the single enabling of $WedgeDetected$: it is a
latency parameter of the detector, not a component of $Connection$.

The talk SUB does \emph{not} survive it, and this is the crux of
biff-9la.  \texttt{safe\_close(nc)} tears down the client on which the
always-on talk subscription was opened, so the SUB is dead.  Yet
$ForceReconnect$ leaves $talkSubGen$ unchanged --- and rightly so:
$generation$ has not advanced yet (it advances only on the subsequent
$EnsureConnected$ dial), so $talkSubGen = generation$ still holds and
the SUB is not yet formally stranded.  The stranding crystallises one
step later, when $EnsureConnected$ dials the replacement client and
bumps $generation$ past $talkSubGen$; only $Subscribe$ repairs it.  A
code path that stops after $EnsureConnected$ --- restoring JetStream and
KV but never re-opening the talk SUB --- is precisely the biff-9la
defect: presence and wall recover, talk does not.

\subsection{Stale Force Teardown (Under-Lock Re-check After Transparent
Reconnect)}

\begin{schema}{StaleForceTeardown}
  \Xi Connection \\
  gateEpoch? : \nat
  \where
  connState = connected \\
  gateEpoch? < reconnectEpoch
\end{schema}

This is the biff-xko fix, and it is the exact analogue of
$StaleRequestTimeout$: where that guards the \emph{redial} race (a
settle on a superseded \emph{generation}), this guards the
\emph{transparent-reconnect} race (a scheduled teardown on a superseded
\emph{epoch}).  The scenario is the DES-042 lineage carried one step
further.  \texttt{\_tracked} decided to force-reconnect a wedged client
and captured, at that gate, the client's in-place reconnect count ---
modelled, exactly as $StaleRequestTimeout$ models $owner?$, as the input
$gateEpoch?$ rather than as stored state.  Between the gate and the lock
nats-py's keepalive completed a transparent reconnect on the \emph{same}
client: $Request(reqTimeout) \to WedgeDetected \to ReconnectSocketUp \to
ReconnectEpochObserved$ carried the connection $connected \to halfOpen
\to reconnecting \to connected$ and, at $ReconnectEpochObserved$, advanced
$reconnectEpoch$ past the value the gate saw.  When
\texttt{\_force\_reconnect} finally takes the lock, the socket it was
dispatched to tear down is healthy again --- and, because the epoch has
been \emph{observed}, this operation ($StaleForceTeardown$) skips.  Had
the lock been taken one step earlier, in the flush window before
$ReconnectEpochObserved$, the epoch would still equal the gate's and the
teardown would proceed: that earlier interleaving is
$FlushWindowForceTeardown$.

The guard $gateEpoch? < reconnectEpoch$ is precisely
\texttt{self.\_reconnect\_epoch != gate\_epoch} on the same client
object: the reconnect count the lock-holder reads has moved past the one
the gate captured, so a transparent reconnect has intervened.  The
operation is a $\Xi Connection$ --- it changes \emph{nothing} --- and
that is the whole of the fix.  The correct behaviour on discovering that
the client has already healed is to \emph{skip the teardown}: leave the
healthy connection exactly as it is.  Modelling the stale re-check as a
no-op from $connected$ makes machine-checked the safety property that
the live connection is invariant under it --- a scheduled force-reconnect
that lands after a transparent reconnect \emph{that has been observed}
cannot move the healthy client to $disconnected$.  The qualification
``observed'' is essential and is the whole subject of
\S\ref{sec:flushwindow}: the guard $gateEpoch? < reconnectEpoch$ can only
recognise a reconnect whose epoch bump has already landed.  A reconnect
that has brought the socket up but whose \texttt{\_on\_reconnect} has not
yet fired ($epochPending = ztrue$, $reconnectEpoch$ still equal to the
gate's) is \emph{not} recognised, and the teardown proceeds --- that is
$FlushWindowForceTeardown$ below.  So there are two transitions that
force-close a client: $ForceReconnect$, from $halfOpen$, a
\emph{genuinely} wedged connection; and $FlushWindowForceTeardown$, from
$connected$, a client whose reconnect is real but not yet observable.
The first is intended.  The second is unavoidable given nats-py's
non-atomic reconnect --- no epoch guard can skip a bump that has not
happened --- and the model's task is not to forbid it but to prove it
benign and self-healing.

The \texttt{is\_connected}-only re-check --- the pre-fix code --- cannot
make this distinction.  \texttt{is\_connected} is $ztrue$ in both
$halfOpen$ and $connected$, the same for a still-wedged socket and a
healed one, so a re-check that reads only \texttt{is\_connected} tears
the client down regardless, charging a $\sim$20\,ms
reconnect-and-reprovision to a connection that needed nothing.  That is
biff-xko.  The frozen variant \texttt{nats-relay-xko-isconnected.tex}
models exactly that re-check: it turns $StaleForceTeardown$ from a $\Xi$
no-op into a $\Delta$ teardown ($connected \to disconnected$) that
flips the observer $spuriousTeardown$, and probcli returns the flip as a
reachable counterexample (\S\ref{sec:modelcheck}).

The precondition is faithful, not merely convenient.
$gateEpoch? < reconnectEpoch$ demands $reconnectEpoch \geq 1$, and a
$connected$ state has $reconnectEpoch \geq 1$ only after a $Reconnect$
--- an actual transparent reconnect.  A freshly-dialled client has
$reconnectEpoch = 0$, so $StaleForceTeardown$ cannot fire against it:
the operation is enabled exactly when a same-object reconnect has really
occurred since the gate, which is the biff-xko precondition and nothing
wider.  Like $StaleRequestTimeout$, it is enabled only from $connected$,
so it adds no transition out of $halfOpen$ (the wedge-liveness proof is
untouched) and removes none (a $connected$ state has many other enabled
operations, so no deadlock is introduced).

The exact guard the code must implement, therefore, has two conjuncts
read under the lock, and $StaleForceTeardown$ together with
$StaleRequestTimeout$ names both.  Proceed with \texttt{safe\_close}
only when \emph{(i)} the client object is still identical
(\texttt{self.\_nc is owner}, the DES-042 guard $StaleRequestTimeout$
already models on $generation$) \emph{and} \emph{(ii)} the client's
reconnect epoch still equals the epoch captured at the gate
(\texttt{self.\_reconnect\_epoch == gate\_epoch}, the new guard this
operation models).  If \emph{either} has advanced --- a full redial
bumped the object and \texttt{\_generation}, or a transparent reconnect
bumped \texttt{\_reconnect\_epoch} on the same object --- the wedge has
already been resolved and the teardown must be skipped.  Concretely, the
relay installs an \texttt{\_on\_reconnect} callback that increments
\texttt{self.\_reconnect\_epoch}; \texttt{\_tracked} records
\texttt{gate\_epoch = self.\_reconnect\_epoch} (and the client identity)
when it schedules the teardown; and \texttt{\_force\_reconnect}, under
the lock, force-closes only if both the identity and the epoch are
unchanged.  Comparing \texttt{\_reconnect\_epoch} is the field the fix
adds; it is the one datum \texttt{is\_connected} does not carry.

One subtlety mirrors the $maxGeneration$ note for $StaleRequestTimeout$:
the $maxEpoch$ saturation clamp is sound only because no stated property
distinguishes reconnects at or above the clamp.  At saturation two
distinct in-place reconnects share the same $reconnectEpoch$, so
$gateEpoch? < reconnectEpoch$ would fail to recognise a reconnect that
completed once the epoch had already pinned to $maxEpoch$ --- benign
here, where the counterexample and the proof both live below the clamp
($gateEpoch? = 0$, $reconnectEpoch = 1$), but load-bearing if the model
is ever strengthened to tie a scheduled teardown to a specific reconnect
rather than to a monotone epoch.  The real \texttt{\_reconnect\_epoch}
is a wide counter that never aliases in practice.

\subsection{Flush Window Force Teardown (Teardown of an Unobserved
Reconnect)}

\begin{schema}{FlushWindowForceTeardown}
  \Delta Connection \\
  gateEpoch? : \nat
  \where
  connState = connected \\
  epochPending = ztrue \\
  gateEpoch? = reconnectEpoch \\
  connState' = disconnected \\
  kvProvisioned' = zfalse \\
  streamsProvisioned' = zfalse \\
  namesProvisioned' = zfalse \\
  wtmpAvailable' = zfalse \\
  generation' = generation \\
  talkSubGen' = talkSubGen \\
  reconnectEpoch' = 0 \\
  spuriousTeardown' = spuriousTeardown \\
  epochPending' = ztrue
\end{schema}

This is the divergence djb traced in wire review, modelled honestly.
$StaleForceTeardown$ is the re-check that finds the reconnect
\emph{observed} ($gateEpoch? < reconnectEpoch$) and skips;
$FlushWindowForceTeardown$ is the same re-check firing one interleaving
earlier, in the flush window, where the reconnect has brought the socket
up ($epochPending = ztrue$) but the epoch bump has not yet landed, so the
lock-holder reads $reconnectEpoch$ still \emph{equal} to the gate's
($gateEpoch? = reconnectEpoch$).  The under-lock guard in
\texttt{\_force\_reconnect} is \texttt{self.\_reconnect\_epoch ==
gate\_epoch} together with \texttt{is\_connected} true and the same
client object; in the flush window every conjunct holds --- the socket is
up, the object is the same, the epoch has not moved --- so the code does
\emph{not} skip.  It calls \texttt{safe\_close} on a client whose socket
is healthy.  The precondition $epochPending = ztrue$ is what distinguishes
this from a freshly-dialled $connected$ client, which also has
$reconnectEpoch = gateEpoch? = 0$ but against which the gate never fired:
the flag marks a reconnect that has completed at the socket and is
awaiting observation, and only such a client can be torn down here.

The effect is the effect of \texttt{\_force\_reconnect}: the client is
discarded, the connection drops to $disconnected$, the handles are
cleared, and $reconnectEpoch$ is reset to $0$ (the relay's
\texttt{self.\_reconnect\_epoch = 0}, line 603).  Two components do
\emph{not} follow $ForceReconnect$'s pattern.  First, $epochPending$ stays
$ztrue$: the deferred \texttt{\_on\_reconnect} scheduled by nats-py is
still pending; \texttt{safe\_close} does not un-schedule it, and it will
land through $ReconnectEpochObserved$ --- now from $disconnected$ --- to
produce the transient $reconnectEpoch = 1 \land connState = disconnected$
that the reconciled recovery property covers.  Second, $spuriousTeardown$
is \emph{not} flipped.  The ghost observer, as defined for
$StaleForceTeardown$, marks the teardown of a \emph{detectably}
reconnected client --- one whose epoch has advanced past the gate.  Here
the reconnect is precisely \emph{not} detectable at teardown time: the
epoch has not moved, and no reading of the client's state
(\texttt{is\_connected}, \texttt{\_reconnect\_epoch}) could tell the
lock-holder to skip.  Marking this as the biff-xko ``spurious teardown of
a detectably-healed client'' would misstate what happened; it is a
distinct, unavoidable event, and the model tracks its aftermath through
$epochPending$ and $reconnectEpoch$ rather than through the observer.
The safety property $AG\,(spuriousTeardown = zfalse)$ thus remains the
faithful statement of what the epoch guard achieves --- it eliminates
every teardown the guard \emph{could} have skipped --- while the residual
flush-window teardown is bounded and self-healing by the separate
liveness property of \S\ref{sec:modelcheck}.

The cost is one unnecessary reconnect-and-reprovision ($\sim$20\,ms, no
data loss), identical to the original biff-xko cost, and the recovery is
the model's existing lazy connect: from $disconnected$, $EnsureConnected$
dials a fresh client, resets the epoch, bumps $generation$ past the stale
callback, and returns to a healthy $connected$ state.  The full path is
$connected \to disconnected \to connected$, closed by operations the
model already contains, so $FlushWindowForceTeardown$ introduces no new
terminal state and no deadlock.  It is enabled only from $connected$, so
it adds no transition out of $halfOpen$ and the biff-tww wedge-liveness
proof is untouched.

\subsection{Stale Request Timeout (Superseded-Generation No-Op)}

\begin{schema}{StaleRequestTimeout}
  \Xi Connection \\
  owner? : \nat
  \where
  connState = connected \\
  owner? < generation
\end{schema}

The formal counterpart of \texttt{\_tracked}'s identity guard
(Copilot: tracked-timeout race).  A JetStream or KV request issued
on generation $owner?$ can still be in flight when its client is
declared dead, force-closed, and replaced: $ForceReconnect$ drops
to $disconnected$ and a fresh $EnsureConnected$ dials a new client,
advancing $generation$ past $owner?$.  When that superseded request
finally times out, the live connection is a \emph{different},
healthy client.  The guard $owner? < generation$ is exactly
\texttt{self.\_nc is not owner}: the settling timeout belongs to a
generation the rebuild has already left behind.

The operation is a $\Xi Connection$ --- it changes nothing.  This
is the whole point: a timeout attributed to a superseded generation
must \emph{not} record a wedge, must \emph{not} move the live
connection to $halfOpen$, and must \emph{not} drive
$ForceReconnect$ of the healthy client.  Without the guard, the
pre-fix code read \texttt{self.\_nc} at exception time and charged
the timeout --- and the force-close --- to whatever client happened
to be current, tearing down a live connection.  Modelling the
superseded timeout as a no-op from $connected$ proves that the live
connection's state is invariant under it: the only transition that
can move $connected \to halfOpen$ is $Request(reqTimeout)$, the
\emph{current}-generation timeout.  The symmetric success case
(\texttt{record\_success} on a superseded client clearing the live
wedge latch) is likewise a no-op, and for the same reason: the
guard fires before any health accounting.  $StaleRequestTimeout$
is enabled only when $generation > 0$ --- a superseded generation
must exist for a stale settle to be possible --- and only from
$connected$, so it neither adds a transition out of $halfOpen$
(the liveness proof is untouched) nor removes one (the current
connection has other enabled operations, so no deadlock is
introduced).

Two subtleties are worth recording.  First, the $maxGeneration = 3$
saturation clamp is sound only because no stated property
distinguishes dials at or above the clamp: at saturation two
distinct dials share the same $generation$ value, so $owner? <
generation$ would treat a superseded timeout on the clamped dial as
\emph{current} --- benign today, but load-bearing if
$StaleRequestTimeout$ is ever strengthened to tie a request to a
specific rebuild rather than to a monotone generation.  Second, the
guard is a partial refinement of \texttt{self.\_nc is not owner}:
$owner? < generation$ captures the redial-to-newer-live case (a
fresh client already dialed), while the client-dropped-to-$None$
-not-yet-redialed case is handled implicitly by the $connState =
connected$ guard --- a settle arriving in that window lands in
$disconnected$, an unmodelled no-op.  ($reconnecting$ is nats-py's
in-place reconnect of the \emph{same} client and generation, so
\texttt{self.\_nc is owner} still holds; it is a current settle, not a
dropped-client window.)

\subsection{Subscribe (Talk SUB Bind / Rebind)}

\begin{schema}{Subscribe}
  \Delta Connection
  \where
  connState = connected \\
  talkSubGen < generation \\
  talkSubGen' = generation \\
  connState' = connState \\
  kvProvisioned' = kvProvisioned \\
  streamsProvisioned' = streamsProvisioned \\
  namesProvisioned' = namesProvisioned \\
  wtmpAvailable' = wtmpAvailable \\
  generation' = generation \\
  reconnectEpoch' = reconnectEpoch \\
  spuriousTeardown' = spuriousTeardown \\
  epochPending' = epochPending
\end{schema}

The poller's re-subscribe transition
(\texttt{poll\_inbox} / \texttt{\_active\_tick}): on each tick, once
the connection is $connected$, the poller compares the generation the
talk SUB is bound to against the current connection generation and, if
they differ, opens a fresh SUB on the live client and records the new
binding.  This one operation covers three cases the code treats
uniformly, because $talkSubGen < generation$ is their common
signature.  The \emph{initial} subscribe: after the first dial,
$talkSubGen = 0 < generation = 1$, and the first tick binds the SUB to
generation $1$.  The \emph{re-subscribe} after a force-reconnect: a
fresh dial has advanced $generation$ past the stranded $talkSubGen$,
and the tick rebinds to the new generation.  The idempotent
\emph{no-op}: when $talkSubGen = generation$ already, the guard fails
and $Subscribe$ is not enabled --- the poller does not open a second,
redundant SUB on a client that already carries one.  The
post-condition $talkSubGen' = generation$ is the whole repair: it moves
the connection from stranded to current in a single step, and it is the
transition the code must implement for the biff-9la fix to hold.  Every
other component of $Connection$ is preserved; a re-subscribe touches
the talk SUB and nothing else.

\subsection{Subscribe Retry (Transient Subscribe Failure)}

\begin{schema}{SubscribeRetry}
  \Xi Connection
  \where
  connState = connected \\
  talkSubGen < generation
\end{schema}

A subscribe attempt can fail: \texttt{subscribe} returns without
establishing a SUB (the JetStream/core request times out, or the client
is momentarily unhealthy), so the binding is not advanced and the next
tick tries again.  $SubscribeRetry$ is the $\Xi Connection$ no-op that
models this outcome --- it is enabled in exactly the stranded state
$Subscribe$ is enabled in, and leaves every component, $talkSubGen$
included, unchanged.  It exists for two reasons.  First, without it the
model would assert that a re-subscribe \emph{always} succeeds on the
first attempt, which the code cannot guarantee; modelling the failure
explicitly is what makes the talk-liveness property a genuine liveness
claim (the SUB is \emph{eventually} rebound) rather than a one-step
safety claim.  Second, it closes the audit: the buggy behaviour ---
retrying forever without ever succeeding --- is now a modelled
transition, so a test partition covers the failure branch rather than
leaving it a $GAP$.  Because $Subscribe$ is enabled alongside it in the
same state, $SubscribeRetry$ adds a stalling path but never removes the
escaping one: the liveness proof ($EF\ talkSubGen = generation$) holds
because \emph{some} path rebinds, even though this operation offers a
path that does not.

%%----------------------------------------------------------------------
\section{Session Presence}
%%----------------------------------------------------------------------

\begin{schema}{SessionStore}
  sessions : SESSIONKEY \pinj KVVALUE \\
  wallValue : KVKEY \pfun KVVALUE
  \where
  \# sessions \leq maxSessions \\
  \# wallValue \leq 1
\end{schema}

Sessions are a partial injection: each session key maps to
exactly one value (the serialised $UserSession$).  $wallValue$
is either empty (no wall) or a single mapping (active wall).
Initially the store is empty: $sessions = \emptyset$ and
$wallValue = \emptyset$.  There is no separate $InitSessionStore$
schema: the model-checked state is the $Connection$ machine (see
\S\ref{sec:modelcheck}), and the session store is exercised by
the operations below only while $connState = connected$.

\subsection{Update Session (Heartbeat / Login)}

\begin{schema}{UpdateSession}
  \Delta SessionStore \\
  \Xi Connection \\
  key? : SESSIONKEY \\
  value? : KVVALUE
  \where
  connState = connected \\
  sessions' = sessions \oplus \{ key? \mapsto value? \} \\
  wallValue' = wallValue \\
  \# sessions' \leq maxSessions
\end{schema}

Creates or refreshes a session.  Override ($\oplus$) ensures
idempotent updates.

\subsection{Delete Session (Logout / TTL Expiry)}

\begin{schema}{DeleteSession}
  \Delta SessionStore \\
  \Xi Connection \\
  key? : SESSIONKEY
  \where
  connState = connected \\
  key? \in \dom sessions \\
  sessions' = \{ key? \} \ndres sessions \\
  wallValue' = wallValue
\end{schema}

Removes a session.  In production, this is triggered by
explicit delete or KV TTL expiry.

\subsection{Set Wall}

\begin{schema}{SetWall}
  \Delta SessionStore \\
  \Xi Connection \\
  wallKey? : KVKEY \\
  wallVal? : KVVALUE
  \where
  connState = connected \\
  sessions' = sessions \\
  wallValue' = \{ wallKey? \mapsto wallVal? \}
\end{schema}

\subsection{Clear Wall}

\begin{schema}{ClearWall}
  \Delta SessionStore \\
  \Xi Connection
  \where
  connState = connected \\
  sessions' = sessions \\
  wallValue' = \emptyset
\end{schema}

%%----------------------------------------------------------------------
\section{KV Watcher State Machine}
%%----------------------------------------------------------------------

This section models the watcher loop from \texttt{\_run\_kv\_watch}
and its supervisor \texttt{\_kv\_watcher\_loop}.

\begin{schema}{Watcher}
  phase : WatcherPhase \\
  shutdownRequested : ZBOOL \\
  restartCount : \nat
  \where
  restartCount \leq 100 \\
  phase = stopped \implies shutdownRequested = ztrue
\end{schema}

The watcher tracks its current phase, whether shutdown has been
requested, and how many times the supervisor has restarted it.
A stopped watcher implies shutdown was requested.  Initially the
watcher is $idle$, with $shutdownRequested = zfalse$ and
$restartCount = 0$.  As with the session store, there is no
separate $InitWatcher$ schema; the watcher operations are enabled
only while $connState = connected$, which is consistent with the
composed invariant $phase \neq idle \implies connState = connected$.

\subsection{Start Watching}

\begin{schema}{StartWatching}
  \Delta Watcher \\
  \Xi Connection
  \where
  connState = connected \\
  phase = idle \\
  shutdownRequested = zfalse \\
  phase' = snapshotting \\
  shutdownRequested' = zfalse \\
  restartCount' = restartCount
\end{schema}

Transitions from idle to snapshotting when the watcher is
created.  The first entries from \texttt{watchall()} are the
snapshot.

\subsection{Snapshot Done}

\begin{schema}{SnapshotDone}
  \Delta Watcher \\
  \Xi Connection
  \where
  connState = connected \\
  phase = snapshotting \\
  shutdownRequested = zfalse \\
  phase' = streaming \\
  shutdownRequested' = zfalse \\
  restartCount' = restartCount
\end{schema}

The $None$ marker from \texttt{updates()} signals the end of
the snapshot.  \textbf{The watcher continues into streaming
phase.}  This is the key property: the old \texttt{async~for}
pattern terminated here (biff-udp).

\subsection{Receive Entry}

\begin{schema}{ReceiveEntry}
  \Delta Watcher \\
  \Xi Connection
  \where
  connState = connected \\
  (phase = snapshotting \lor phase = streaming) \\
  shutdownRequested = zfalse \\
  phase' = phase \\
  shutdownRequested' = zfalse \\
  restartCount' = restartCount
\end{schema}

A KV entry is received.  The phase does not change --- entries
arrive during both snapshotting and streaming.

\subsection{Handle Timeout}

\begin{schema}{HandleTimeout}
  \Delta Watcher \\
  \Xi Connection
  \where
  connState = connected \\
  phase = streaming \\
  shutdownRequested = zfalse \\
  phase' = streaming \\
  shutdownRequested' = zfalse \\
  restartCount' = restartCount
\end{schema}

A \texttt{TimeoutError} from \texttt{updates(timeout=5.0)} is
caught and the loop continues.  \textbf{The watcher stays in
streaming phase.}  Without the catch, this would propagate to
the supervisor and trigger a restart.

\subsection{Handle Error (Supervisor Restart)}

\begin{schema}{HandleError}
  \Delta Watcher \\
  \Xi Connection
  \where
  connState = connected \\
  phase \neq idle \\
  phase \neq stopped \\
  shutdownRequested = zfalse \\
  phase' = snapshotting \\
  shutdownRequested' = zfalse \\
  restartCount' = restartCount + 1
\end{schema}

An unrecoverable error causes the supervisor to restart the
watcher.  The new watcher begins in snapshotting phase
(a fresh \texttt{watchall()}).

\subsection{Request Shutdown}

\begin{schema}{RequestShutdown}
  \Delta Watcher \\
  \Xi Connection
  \where
  connState = connected \\
  shutdownRequested = zfalse \\
  shutdownRequested' = ztrue \\
  phase' = stopped \\
  restartCount' = restartCount
\end{schema}

The shutdown event signals the watcher to exit.  The loop
checks \texttt{shutdown.is\_set()} on every iteration.

%%----------------------------------------------------------------------
\section{Composed System State}
%%----------------------------------------------------------------------

\begin{schema}{NatsRelayState}
  Connection \\
  SessionStore \\
  Watcher
  \where
  phase \neq idle \implies connState = connected \\
  phase = streaming \implies kvProvisioned = ztrue
\end{schema}

The watcher can only be active when the connection is
established.  Streaming implies KV is provisioned (the watcher
reads from it).  With the generalised handle-validity invariant,
$phase = streaming \implies kvProvisioned = ztrue$ also implies
$phase = streaming \implies connState = connected$: a wedge
($halfOpen$, where $kvProvisioned = zfalse$) cannot coexist with a
streaming watcher, which is consistent --- a wedge manifests to
the watcher as a $weError$ and drives it through $HandleError$.

The composed initial state is: $connState = disconnected$ with no
handles and no talk subscription ($talkSubGen = 0$, $InitConnection$),
an empty session store, and an $idle$ watcher.  $InitConnection$ is the sole initialisation schema so
that the model checker explores the $Connection$ lifecycle as a
single, well-posed state machine; the session and watcher
components are exercised as connection-gated operations rather
than as independent initial states (see \S\ref{sec:modelcheck}).

%%----------------------------------------------------------------------
\section{Key Properties and Model Checking}
\label{sec:modelcheck}
%%----------------------------------------------------------------------

The specification captures these properties, each of which can be
verified by model checking with probcli (\texttt{z-spec test}):

\begin{enumerate}
  \item \textbf{Handle validity}: valid JetStream, sessions-KV, and
    names-KV handles exist if and only if $connState = connected$.
    A $halfOpen$, $disconnected$, $reconnecting$, or $closed$
    connection has no valid handles.  This generalises DES-029 and
    is the safety half of the biff-tww fix: stale handles cached
    during a wedge are, in the model, no handles at all.

  \item \textbf{Wedge liveness (biff-tww)}: no reachable state is
    stuck in $halfOpen$.  Two independent recovery paths leave the
    wedge: the keepalive path $WedgeDetected \to ReconnectSocketUp$ (via
    $reconnecting$, on the same client) and the proactive path
    $ForceReconnect \to EnsureConnected$ (via $disconnected$, on a
    fresh client).  Each is always enabled in $halfOpen$, so the
    connection cannot deadlock in the wedge.  The proof is by
    deadlock analysis: removing \emph{every} operation enabled in
    $halfOpen$ (the pre-fix code, whose keepalive ping never fired
    in time, with no proactive detector) makes $halfOpen$ a
    reachable terminal state.  probcli surfaces it as a deadlock
    along the trace $EnsureConnected \to Request(reqTimeout)$,
    which \emph{is} biff-tww.  With either recovery transition
    present the state space is deadlock-free.

  \item \textbf{Generation-guarded wedge (tracked-timeout race)}: a
    timeout on a \emph{superseded} generation
    ($StaleRequestTimeout$, $owner? < generation$) is a no-op ---
    the live $connected$ connection is invariant under it.  Only
    the \emph{current}-generation timeout $Request(reqTimeout)$
    moves $connected \to halfOpen$, so a settle on a client the
    rebuild has already replaced can never drive $ForceReconnect$
    of the healthy connection.  $generation$ advances only on a
    fresh dial ($EnsureConnected$ from $disconnected$), never on
    nats-py's in-place $Reconnect$, mirroring the relay's
    \texttt{\_generation} token.

  \item \textbf{Epoch-guarded teardown (biff-xko)}: a scheduled
    force-reconnect whose gate observed the epoch $gateEpoch?$ is a no-op
    against a client that has since transparently reconnected
    ($StaleForceTeardown$, $gateEpoch? < reconnectEpoch$) --- the live
    $connected$ connection is invariant under it.  The safety property is
    stated over reachable states as the observer-guarded formula

    \begin{quote}
      $AG\,(\,spuriousTeardown = zfalse\,)$
    \end{quote}

    where $spuriousTeardown$ is a ghost flipped only by a teardown of a
    detectably-reconnected client.  probcli reports this \textbf{TRUE}
    for the shipped spec: the epoch-guarded re-check never tears down a
    healed connection, because $reconnectEpoch$ advances on nats-py's
    in-place $Reconnect$ where \texttt{is\_connected} cannot distinguish
    healed from wedged.  It advances \emph{independently} of $generation$
    --- a transparent reconnect bumps $reconnectEpoch$ and leaves
    $generation$ fixed, a fresh dial the reverse --- so the pair
    $(generation, reconnectEpoch)$ is the full transport identity the
    two guards $StaleRequestTimeout$ and $StaleForceTeardown$ compare
    against.  The pre-fix defect is reproduced in the frozen variant
    \texttt{nats-relay-xko-isconnected.tex}, whose re-check reads only
    \texttt{is\_connected} and so turns $StaleForceTeardown$ into a
    $connected \to disconnected$ teardown: the same formula then goes
    \textbf{FALSE}, with probcli returning the eight-step trace
    $EnsureConnected \to Subscribe \to Request(reqTimeout) \to
    WedgeDetected \to Reconnect \to StaleForceTeardown$ --- a client
    force-closed one step after nats-py had transparently reconnected it.
    That trace \emph{is} biff-xko.  Liveness is not regressed: the
    genuinely-wedged teardown $ForceReconnect$ (enabled from $halfOpen$)
    is untouched, and probcli confirms $AG\,(connState = halfOpen
    \Rightarrow EF\;connState = disconnected)$ remains \textbf{TRUE}.
    In the shipped spec the atomic $Reconnect$ this trace names is split
    into $ReconnectSocketUp \to ReconnectEpochObserved$; the frozen
    variant is unchanged and retains the single-step $Reconnect$, so its
    counterexample trace stands verbatim.

  \item \textbf{Flush-window teardown is bounded and self-healing
    (biff-xko, wire-review correction)}: the epoch guard of the previous
    property recognises only a reconnect whose epoch bump has \emph{landed}.
    nats-py's reconnect is not atomic --- socket up, then flush, then the
    epoch-bumping \texttt{\_on\_reconnect} --- so the model splits it into
    $ReconnectSocketUp$ (socket $connected$, $epochPending = ztrue$,
    $reconnectEpoch$ unchanged) and $ReconnectEpochObserved$ (the bump
    lands).  Between them, $FlushWindowForceTeardown$ (guard $connState =
    connected \land epochPending = ztrue \land gateEpoch? = reconnectEpoch$)
    force-closes a client whose socket is healthy but whose reconnect is
    not yet observable --- the guard \texttt{\_reconnect\_epoch ==
    gate\_epoch} passes because the epoch has not moved.  This is the
    divergence djb traced: the state it reaches, $connState = disconnected
    \land reconnectEpoch = 1$ (the deferred bump landing after the teardown
    reset the epoch), \emph{violates} the old state invariant $connState
    \in \{ disconnected, closed \} \implies reconnectEpoch = 0$.  That
    invariant is therefore reclassified from a global always-property to an
    eventually-true recovery property, proved as

    \begin{quote}
      $AG\,(\,connState = disconnected \land reconnectEpoch > 0
      \;\Rightarrow\; EF\;reconnectEpoch = 0\,)$
    \end{quote}

    which probcli reports \textbf{TRUE}: the next $EnsureConnected$ dial
    resets the epoch.  The state $connState = disconnected \land
    reconnectEpoch > 0$ is confirmed \emph{reachable}
    ($EF$ \textbf{TRUE}), so the property is not vacuous --- the model
    genuinely reaches the state the old invariant forbade and proves it
    transient.  The flush-window teardown itself is bounded and
    self-healing:

    \begin{quote}
      $AG\,(\,connState = disconnected \land epochPending = ztrue
      \;\Rightarrow\; EF\;(connState = connected \land epochPending =
      zfalse)\,)$
    \end{quote}

    is \textbf{TRUE} --- every trace through the teardown returns to a
    healthy $connected$ state via a fresh dial.  The observer property
    $AG\,(spuriousTeardown = zfalse)$ is unaffected: the flush-window
    teardown does not flip $spuriousTeardown$, because the reconnect it
    tears down is by definition \emph{not} detectable at teardown time
    (the epoch has not advanced), so it is a distinct event from the
    detectably-healed teardown the observer marks.  probcli confirms the
    biff-tww wedge-liveness and biff-9la talk-liveness properties both
    remain \textbf{TRUE} across the extended state space, and the
    constraint-based deadlock check finds no deadlock: the new operations
    add no terminal state.

  \item \textbf{Talk-subscription liveness (biff-9la)}: the always-on
    talk SUB is tracked by $talkSubGen$, the generation it is bound to,
    with the state invariant $talkSubGen \leq generation$ (a SUB is
    never bound to a client never dialled).  A live connection whose SUB
    is stranded on a superseded generation always eventually re-binds it:

    \begin{quote}
      $AG\,(\,connState \neq closed \land talkSubGen < generation
      \;\Rightarrow\; EF\;talkSubGen = generation\,)$
    \end{quote}

    probcli reports this CTL formula \textbf{TRUE} (the restriction to
    $connState = connected$ is TRUE as well; only the terminal $closed$
    state, in which the connection is permanently shut down and carries
    no talk obligation, is excluded).  The escape is $Subscribe$
    (guard $talkSubGen < generation$, effect $talkSubGen' = generation$),
    the poller's generation-tracked re-subscribe; it is preserved
    unchanged across nats-py's in-place $Reconnect$ (same client, same
    generation, SUB replayed) and re-established after $ForceReconnect$
    once the replacement dial advances $generation$ past $talkSubGen$.
    The property is a genuine liveness claim, not a one-step safety
    claim: $SubscribeRetry$ models a subscribe that fails and is retried
    next tick, so the SUB is \emph{eventually}, not \emph{immediately},
    re-bound.  The pre-fix defect is reproduced by replacing the
    $Subscribe$ guard with the ``re-subscribe only when the handle is
    \texttt{None}'' condition $talkSubGen = 0$: the same CTL formula
    then goes \textbf{FALSE}, with probcli returning the counterexample
    $connState = connected \land generation = 3 \land talkSubGen = 2$
    --- a live client whose talk SUB is stuck on the superseded
    generation $2$, never re-opened.  That state \emph{is} biff-9la.

  \item \textbf{Snapshot survival}: $SnapshotDone$ transitions to
    $streaming$, never to $idle$ or $stopped$.  The watcher
    remains alive.

  \item \textbf{Timeout resilience}: $HandleTimeout$ keeps the
    watcher in $streaming$ phase.  No restart, no phase change.

  \item \textbf{Error restart}: Only $HandleError$ increments
    $restartCount$.  Snapshot-done and timeout do not.

  \item \textbf{Connection prerequisite}: The watcher cannot be
    active ($phase \neq idle$) when the connection is not
    $connected$; every session and watcher operation is enabled
    only while $connState = connected$.

  \item \textbf{Session uniqueness}: The partial injection
    ($\pinj$) on sessions guarantees each session key maps to
    exactly one value.
\end{enumerate}

\end{document}
