%
% notification.tex
%
% Z Specification for Biff Notification System
% Formal model of NATS connection, MCP server, filesystem, and
% status bar states and transitions in the send and receive flow.
%
% Corrected model: eager session capture during server lifespan
% ensures the suspenders path always has a valid session reference.
%

\documentclass[a4paper,10pt,fleqn]{article}
\usepackage[margin=1in]{geometry}
\usepackage{fuzz}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\begin{document}

\title{Biff Notification System: A Z Specification}
\author{Formal Model of Push Notification State Transitions}
\date{March 2026}
\hypersetup{
  pdftitle={Biff Notification System: A Z Specification},
  pdfauthor={punt-labs},
  pdfsubject={Z Specification},
  pdfcreator={fuzz/probcli}
}
\maketitle

\tableofcontents
\newpage

%%----------------------------------------------------------------------
\section{Introduction}
%%----------------------------------------------------------------------

Biff's push notification system delivers real-time updates (wall
broadcasts and talk messages) to Claude~Code's status bar through two
independent channels: the MCP protocol (fast, tool-list-changed
notification) and a filesystem status file (slow, polled by the status
bar command).

The notification pipeline has a ``belt and suspenders'' design: tool
handlers fire notifications via the MCP request context (belt path),
while a background poller fires them via a stored
\texttt{ServerSession} reference (suspenders path).  NATS subscription
callbacks and KV watchers only wake the poller from napping; the
poller detects the changes on its next active tick and fires the
notification from the MCP server's own coroutine context.

The session reference is captured eagerly during the server's lifespan
startup, before the NATS connection is established and before the
background poller begins.  This ensures the suspenders path always has
a valid session reference whenever it needs to fire a notification.

The key invariant is: \emph{a connected NATS client implies a captured
session}.  Since NATS subscription is the mechanism that delivers
external talk messages, and since the session is captured before NATS
connects, the poller always has a valid notification path.  NATS
callbacks and KV watchers do not fire notifications directly---they
are unreliable from background coroutine contexts---but they wake the
poller from napping, so the next active tick (2~seconds) detects the
changes via its own reads and fires the notification.

This specification models the four state spaces---NATS connection,
MCP server session, filesystem, and status bar---and the operations
that transition them during the wall and talk notification flows.

A second delivery mechanism sits alongside the status bar: the server
surfaces incoming activity by mutating a \emph{tool description} and
firing \texttt{tools/list\_changed}.  The \texttt{read\_messages}
description gains \texttt{(N unread)}; the \texttt{talk} description gains
a \texttt{[TALK] \ldots} banner with an accept hint.
Sections~\ref{sec:marker} to~\ref{sec:markerfix2} model these two
channels and establish the governing invariant: for each channel the
marker is present exactly when its activity exists, and a talk marker
always carries a session-naming accept hint.  The read channel is
self-clearing and serves as the reference; the talk channel, as
originally built (biff-9la), was not.  The stale-marker defect is
reproduced here as a reachable ProB state and then proved unreachable
once the reconciling transitions of Section~\ref{sec:markerfix} are in
place.

%%----------------------------------------------------------------------
\section{Basic Types}
%%----------------------------------------------------------------------

\begin{zed}
  [USER, MSG, SOURCEKEY, SESSIONKEY]
\end{zed}

$USER$ identifies a biff user.  $MSG$ is an opaque message body.
$SOURCEKEY$ uniquely identifies a display queue item (e.g.\
\texttt{wall:2026-03-01T12:00:00} or \texttt{talk:eric}).
$SESSIONKEY$ is a composite \texttt{user:tty} key naming one specific
session; it is the same notion as the $SESSIONKEY$ of $talk.tex$, and it
is what an accept must target (\texttt{talk @user:tty}).  A pending talk
invite that retains its inviter's $SESSIONKEY$ can render an accept hint
that names a session; one that has dropped it can only render a bare
\texttt{@user}, which is not a runnable accept command.

%%----------------------------------------------------------------------
\section{Free Types}
%%----------------------------------------------------------------------

\begin{zed}
  ZBOOL ::= ztrue | zfalse
\end{zed}

\begin{zed}
  NatsState ::= natsDisconnected | natsConnected | natsClosed
\end{zed}

The NATS client connection lifecycle: initially disconnected, lazily
connected on first use, and permanently closed on shutdown.

\begin{zed}
  SessionState ::= sessNone | sessCaptured
\end{zed}

The MCP \texttt{ServerSession} reference: either not yet captured
(\texttt{sessNone}) or captured during server lifespan startup
(\texttt{sessCaptured}).  Once captured, the session reference
is never invalidated.

\begin{zed}
  ItemKind ::= itemWall | itemTalk
\end{zed}

Display queue items are either persistent wall announcements or
ephemeral talk messages.

\begin{zed}
  NotifyResult ::= notifySent | notifyDropped
\end{zed}

The outcome of a \texttt{notify\_tool\_list\_changed} attempt.
In the corrected system, $notifyDropped$ is only the initial
default value; all notification-triggering operations produce
$notifySent$.

%%----------------------------------------------------------------------
\section{Global Constants}
%%----------------------------------------------------------------------

The display queue and the unread counter are bounded.  In the running
system the display rotation queue holds at most twenty items and the
unread count is capped at one hundred.  Pinning those concrete numbers
into the model, however, makes exhaustive checking hopeless: a sequence
of up to twenty source keys over even two carriers, crossed with an
unread count ranging over $0 \upto 100$, yields hundreds of thousands
of reachable states before any interesting property is reached.

Following the convention of the sister specification $talk.tex$, the
two bounds are therefore not hard-wired.  Each is the cardinality of an
abstract carrier---$QSLOT$ for queue slots, $USLOT$ for unread slots---so
the bound is a \emph{scoping parameter} ProB shrinks for exhaustive
checking rather than a fixed value.  A third carrier, $ASLOT$, plays the
same role for the pending-invite time-to-live: $maxInviteAge$ is its
cardinality, and a pending invite whose age reaches that bound may be
aged out.  All three carriers are required non-empty, so each bound is at
least one; at $\# ASLOT = 1$ the clock has a single tick, which is the
smallest model in which a fresh invite (age $0$) and an expired one (age
$1$) are distinct.

\begin{zed}
  [QSLOT, USLOT, ASLOT]
\end{zed}

\begin{axdef}
  maxQueueSize : \nat \\
  maxUnreadCount : \nat \\
  maxInviteAge : \nat
  \where
  QSLOT \neq \emptyset \\
  USLOT \neq \emptyset \\
  ASLOT \neq \emptyset \\
  maxQueueSize = \# QSLOT \\
  maxUnreadCount = \# USLOT \\
  maxInviteAge = \# ASLOT
\end{axdef}

Default descriptions and the idle marker are opaque constants:

\begin{axdef}
  wallDescDefault : MSG \\
  talkDescDefault : MSG \\
  readDescDefault : MSG \\
  idleMarker : MSG
\end{axdef}

$readDescDefault$ is the base \texttt{read\_messages} tool description,
carrying no unread marker.  Its role in the read channel mirrors that of
$talkDescDefault$ in the talk channel and $wallDescDefault$ in the wall
channel.

The intended verification is exhaustive model checking at set size~2,
which supplies two distinct carrier elements throughout.  Modelling the
tool-context brackets (Section~\ref{sec:toolcontext}) roughly doubles the
reachable state count and, by making the belt path live, opens the
receive-flow subspace behind $TalkStart$; at the full queue and unread
bounds the check no longer terminates in a practical time.  The bounded
carriers exist precisely for this: shrinking $QSLOT$, $USLOT$ and $ASLOT$
to one keeps the exploration complete while retaining every distinction
the properties depend on.

\begin{verbatim}
probcli docs/notification.tex -model_check -noltl -coverage \
  -p DEFAULT_SETSIZE 2 \
  -card SOURCEKEY 1 -card SESSIONKEY 1 \
  -card QSLOT 1 -card USLOT 1 -card ASLOT 1 \
  -p MAX_INITIALISATIONS 16 -p MAX_OPERATIONS 8000
\end{verbatim}

Each carrier may be shrunk independently with \texttt{-card}~$X$~$n$
without disturbing the message and key carriers.  $MAX\_INITIALISATIONS$
must exceed the number of identities $Init$ leaves unconstrained, or ProB
silently drops initial states and reports an incomplete exploration; at
these bounds $Init$ is fully determined, so exploration reports
$INITIALISATION$:$8$ against $SETUP\_CONSTANTS$:$8$ with the bound never
binding, and terminates with ``all open states visited''---a complete
check, not a truncated one.

%%----------------------------------------------------------------------
\section{State}
\label{sec:state}
%%----------------------------------------------------------------------

The system state is flattened into a single schema for ProB
compatibility.  It covers four subsystems: NATS connection state,
MCP session capture, the display queue, the filesystem status file,
the status bar, and the last notification outcome.

\begin{schema}{NotificationState}
  natsConn : NatsState \\
  kvProvisioned : ZBOOL \\
  talkSubscribed : ZBOOL \\
  session : SessionState \\
  toolContext : ZBOOL \\
  wallDesc : MSG \\
  talkDesc : MSG \\
  readDesc : MSG \\
  wallDescChanged : ZBOOL \\
  talkDescChanged : ZBOOL \\
  unreadCount : \nat \\
  talkPending : USER \pfun SESSIONKEY \\
  talkBare : \power USER \\
  talkQueued : \nat \\
  talkConnected : ZBOOL \\
  pendingAge : \nat \\
  queueItems : \seq SOURCEKEY \\
  queueKinds : SOURCEKEY \pfun ItemKind \\
  queueTexts : SOURCEKEY \pfun MSG \\
  currentIndex : \nat \\
  fileExists : ZBOOL \\
  fileUnreadCount : \nat \\
  fileItems : \seq SOURCEKEY \\
  barLine2 : MSG \\
  barStale : ZBOOL \\
  lastNotifyResult : NotifyResult \\
  staleWithdrawHit : ZBOOL
  \where
  \dom queueKinds = \ran queueItems \\
  \dom queueTexts = \ran queueItems \\
  \# queueItems \leq maxQueueSize \\
  (queueItems \neq \langle \rangle \implies
  currentIndex < \# queueItems) \\
  (queueItems = \langle \rangle \implies
  currentIndex = 0) \\
  fileUnreadCount \leq maxUnreadCount \\
  unreadCount \leq maxUnreadCount \\
  talkQueued \leq maxQueueSize \\
  pendingAge \leq maxInviteAge \\
  ((\dom talkPending = \emptyset \land talkBare = \emptyset) \implies
  pendingAge = 0) \\
  \dom talkPending \cap talkBare = \emptyset \\
  ((talkDesc \neq talkDescDefault) \iff \\
  \quad (\dom talkPending \neq \emptyset \lor talkBare \neq \emptyset \lor \\
  \quad~ talkQueued > 0 \lor talkConnected = ztrue)) \\
  ((readDesc \neq readDescDefault) \iff unreadCount > 0) \\
  (talkDesc \neq talkDescDefault \implies talkBare = \emptyset) \\
  (talkSubscribed = ztrue \implies
  natsConn = natsConnected) \\
  (kvProvisioned = ztrue \implies
  natsConn = natsConnected) \\
  (natsConn = natsConnected \implies
  session = sessCaptured) \\
  staleWithdrawHit = zfalse
\end{schema}

Key invariants: queue metadata maps have the same domain as the
queue sequence range; subscription and KV require a live connection;
the current index is always within bounds.  The critical invariant
is the last: \emph{a connected NATS client implies a captured session}.
This ensures the suspenders notification path (the background poller)
is always available when description-changed flags are set.

\paragraph{The tool-description channels.}  Three of the components above
model tool descriptions that Claude~Code shows to the agent, and two more
model the underlying activity that a description marker is supposed to
track.  These are the subject of the correctness argument in
Section~\ref{sec:marker}.

$readDesc$ is the \texttt{read\_messages} tool description; it gains a
\texttt{(N unread)} marker when mail is pending.  Its source of truth is
$unreadCount$, the server-side count of unread messages, which
\emph{decreases} as mail is read---so a description tied to it clears
itself.  This is the reference channel: it is known to behave correctly,
and it is modelled here so that the talk channel can be judged against a
worked example rather than against prose.

$talkDesc$ is the \texttt{talk} tool description; it gains a
\texttt{[TALK] \ldots} marker, including an accept hint, when a talk
invite, queued message, or live connection exists.  Its activity has
three sources: $talkPending$, a partial function mapping each user who
has invited us to that invite's originating $SESSIONKEY$; $talkQueued$,
the count of undrained talk messages; and $talkConnected$, whether a
conversation is live.  A pending invite in $talkPending$ retains the
inviter's session key, so its accept hint can name a session.

$talkBare$ is the pathological representation: a set of users whose
pending invite was recorded \emph{without} a session key---the tty was
dropped.  A user in $talkBare$ is genuine activity (the marker appears),
but its accept hint can only read \texttt{talk @user} with no
\texttt{:tty}, which is not a runnable accept command.  The structural
invariant $\dom talkPending \cap talkBare = \emptyset$ says a single
invite is recorded one way or the other, never both.  In a correct
implementation $talkBare$ is empty in every reachable state; the model
admits it precisely so the defect can be reached and then proved
unreachable.

\paragraph{The pending-invite clock.}  $pendingAge$ is an abstract clock
for the pending set, bounded by $maxInviteAge$.  It advances on poller
ticks ($AgeTick$) while any invite is pending and is pinned to zero
whenever no invite is pending---the invariant
$(\dom talkPending = \emptyset \land talkBare = \emptyset) \implies
pendingAge = 0$ says exactly this, so a fresh invite always starts its
life at age zero.  Its sole purpose is to make the time-to-live path of
Section~\ref{sec:markerfix} reachable: an invite may be expired
($ExpirePendingInvite$) only once the clock has reached its bound.  The
clock is deliberately global rather than per-invite; at these bounds that
loses nothing the marker property depends on, and it keeps the state
factor at two rather than one per user.

%%----------------------------------------------------------------------
\section{Initialization}
%%----------------------------------------------------------------------

\begin{schema}{Init}
  NotificationState'
  \where
  natsConn' = natsDisconnected \\
  kvProvisioned' = zfalse \\
  talkSubscribed' = zfalse \\
  session' = sessNone \\
  toolContext' = zfalse \\
  wallDesc' = wallDescDefault \\
  talkDesc' = talkDescDefault \\
  readDesc' = readDescDefault \\
  wallDescChanged' = zfalse \\
  talkDescChanged' = zfalse \\
  unreadCount' = 0 \\
  talkPending' = \emptyset \\
  talkBare' = \emptyset \\
  talkQueued' = 0 \\
  talkConnected' = zfalse \\
  pendingAge' = 0 \\
  queueItems' = \langle \rangle \\
  queueKinds' = \emptyset \\
  queueTexts' = \emptyset \\
  currentIndex' = 0 \\
  fileExists' = zfalse \\
  fileUnreadCount' = 0 \\
  fileItems' = \langle \rangle \\
  barLine2' = idleMarker \\
  barStale' = ztrue \\
  lastNotifyResult' = notifyDropped \\
  staleWithdrawHit' = zfalse
\end{schema}

The initial state has $session' = sessNone$ and
$natsConn' = natsDisconnected$.  The invariant
$natsConn = natsConnected \implies session = sessCaptured$
holds vacuously because $natsConn \neq natsConnected$.

%%----------------------------------------------------------------------
\section{Session Capture}
%%----------------------------------------------------------------------

The MCP server lifespan startup captures the session reference
eagerly, before the NATS connection is established and before the
background poller begins.  This is the key operation that prevents
the biff-8g0 bug class: by the time any NATS callback can fire,
the session is guaranteed to be captured.

\begin{schema}{CaptureSession}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  session = sessNone \\
  natsConn = natsDisconnected \\
  session' = sessCaptured \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

After $CaptureSession$, the system has $session = sessCaptured$
with $natsConn = natsDisconnected$.  The invariant holds, and
$EnsureConnected$ can now be applied---the resulting post-state
$natsConn' = natsConnected \land session' = sessCaptured$ satisfies
the invariant.

%%----------------------------------------------------------------------
\section{NATS Connection Operations}
%%----------------------------------------------------------------------

\subsection{Ensure Connected}

Lazy connection: transitions from disconnected to connected,
provisioning KV and stream infrastructure.  Idempotent when
already connected.

The state invariant
$natsConn = natsConnected \implies session = sessCaptured$
creates an implicit precondition: since this operation sets
$natsConn' = natsConnected$ and preserves $session' = session$,
the post-state requires $session = sessCaptured$.  The system
must execute $CaptureSession$ before $EnsureConnected$.

\begin{schema}{EnsureConnected}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  natsConn = natsDisconnected \lor natsConn = natsConnected \\
  natsConn' = natsConnected \\
  kvProvisioned' = ztrue \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Close Connection}

Permanent shutdown.  Clears all NATS-dependent state.  The session
reference is preserved---it is not invalidated by disconnection.
The invariant holds because $natsConn' = natsClosed$, so the
antecedent $natsConn = natsConnected$ is false.

\begin{schema}{CloseConnection}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  natsConn' = natsClosed \\
  kvProvisioned' = zfalse \\
  talkSubscribed' = zfalse \\
  session' = session \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

%%----------------------------------------------------------------------
\section{Tool Context}
\label{sec:toolcontext}
%%----------------------------------------------------------------------

A tool handler runs \emph{inside} the MCP request-response cycle; the
background poller and the NATS subscription callbacks run \emph{outside}
it.  The component $toolContext$ records which of the two the system is
currently in: $ztrue$ during a tool call, $zfalse$ otherwise.  The
distinction is not decoration.  It selects the notification path---belt
when a request context is available, suspenders when it is not---and it
gates every operation that can only occur inside a handler ($WallPost$,
$TalkStart$, $NotifyBelt$) against every operation that can only occur
in the background ($NotifySuspenders$, $KVWallReceive$,
$NatsTalkCallback$, $PollTick$).

The earlier drafts left $toolContext$ at its initial $zfalse$ with no
operation to raise it.  The consequence was quiet but total: no reachable
state had $toolContext = ztrue$, so the belt-path operations were enabled
in no reachable state, and the guarantees stated for them
($NotifyGuarantee$, $WallNotifySuccess$) held only \emph{vacuously}.  A
vacuous proof is not a proof of anything.  The remedy is to model the
entry into and exit from a tool handler explicitly, so that the belt path
is reachable and its obligations bite.

$EnterToolContext$ marks the start of a tool call; $LeaveToolContext$
marks its end.  Each toggles $toolContext$ and changes nothing else, so
neither can disturb any invariant.  Their guards make the two mutually
exclusive: one may enter only from the background, and leave only from a
handler.

\begin{schema}{EnterToolContext}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  toolContext = zfalse \\
  toolContext' = ztrue \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\begin{schema}{LeaveToolContext}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  toolContext = ztrue \\
  toolContext' = zfalse \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

A tool handler is therefore modelled as $EnterToolContext$, followed by
the handler's own operation on the belt path, followed by
$LeaveToolContext$; background work happens between these brackets, never
inside them.  This is the smallest change that makes the belt path
reachable without weakening any existing guard.

%%----------------------------------------------------------------------
\section{Notification Core}
%%----------------------------------------------------------------------

This is the central mechanism.  Two paths send the MCP
\texttt{tools/list\_changed} notification.  Both always succeed.

\subsection{Belt Path (Tool Context Available)}

Inside a tool handler: notification piggybacks on the MCP response.
This is the preferred path when tool context is available because
it guarantees synchronous delivery within the request-response cycle.

\begin{schema}{NotifyBelt}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  toolContext = ztrue \\
  lastNotifyResult' = notifySent \\
  session' = session \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = zfalse \\
  talkDescChanged' = zfalse \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = zfalse \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Suspenders Path (No Tool Context)}

Background poller: no request context, but the eagerly captured
session reference is used to send the notification.  NATS callbacks
do not use this path directly---they set description-changed flags
that the poller detects on its next tick.  The state invariant
guarantees $session = sessCaptured$ whenever NATS is connected,
so the poller always succeeds when needed.

\begin{schema}{NotifySuspenders}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  toolContext = zfalse \\
  session = sessCaptured \\
  lastNotifyResult' = notifySent \\
  session' = session \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = zfalse \\
  talkDescChanged' = zfalse \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = zfalse \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Total Notification}

The complete notification operation is the disjunction of the two
paths.  Exactly one fires: $NotifyBelt$ when $toolContext = ztrue$,
$NotifySuspenders$ when $toolContext = zfalse$.

Because the state invariant guarantees $session = sessCaptured$
whenever NATS is connected, and because all notification-triggering
operations require a NATS connection, $NotifySuspenders$ is always
applicable when $toolContext = zfalse$ and a description has changed.
No notification is ever dropped---NATS callbacks defer to the poller
and wake it from napping, so the next active tick fires the notification.

%%----------------------------------------------------------------------
\section{Send Flow Operations}
%%----------------------------------------------------------------------

\subsection{Wall Post (Tool Handler)}

The wall tool executes inside a tool handler context.  It
mutates the wall description, adds a display item, and fires
notification via the belt path.  The precondition requires
tool context and a live NATS connection.

\begin{schema}{WallPost}
  \Delta NotificationState \\
  wallText? : MSG \\
  wallKey? : SOURCEKEY
  \where
  pendingAge' = pendingAge \\
  toolContext = ztrue \\
  natsConn = natsConnected \\
  wallDesc' \neq wallDesc \\
  wallDescChanged' = ztrue \\
  (wallKey? \notin \ran queueItems \implies \\
  \quad~ queueItems' = queueItems \cat \langle wallKey? \rangle \land \\
  \quad~ queueKinds' = queueKinds \cup \\
  \quad~ \quad~ \{ wallKey? \mapsto itemWall \} \land \\
  \quad~ queueTexts' = queueTexts \cup \\
  \quad~ \quad~ \{ wallKey? \mapsto wallText? \}) \\
  (wallKey? \in \ran queueItems \implies \\
  \quad~ queueItems' = queueItems \land \\
  \quad~ queueKinds' = queueKinds \land \\
  \quad~ queueTexts' = queueTexts) \\
  session' = session \\
  lastNotifyResult' = notifySent \\
  barStale' = zfalse \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  talkDesc' = talkDesc \\
  talkDescChanged' = zfalse \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Talk Start (Tool Handler)}

The talk tool initiates a conversation.  Runs inside tool context,
and arranges for the background poller to subscribe to NATS talk
notifications.

\begin{schema}{TalkStart}
  \Delta NotificationState \\
  partner? : USER
  \where
  pendingAge' = pendingAge \\
  toolContext = ztrue \\
  natsConn = natsConnected \\
  session' = session \\
  lastNotifyResult' = notifySent \\
  talkSubscribed' = ztrue \\
  barStale' = zfalse \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  wallDescChanged' = zfalse \\
  talkDesc' = talkDesc \\
  talkDescChanged' = zfalse \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{External Message Deliver (CLI)}

A CLI client or another MCP server delivers a message via NATS.
This publishes to JetStream (reliable) and then publishes a
talk notification on core NATS (fire-and-forget).  No tool
context exists on the sender side.  The receiver's $NatsTalkCallback$
handles the receive-side effects.

\begin{schema}{ExternalDeliver}
  \Delta NotificationState \\
  body? : MSG \\
  toUser? : USER
  \where
  pendingAge' = pendingAge \\
  natsConn = natsConnected \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

%%----------------------------------------------------------------------
\section{Receive Flow Operations}
%%----------------------------------------------------------------------

\subsection{KV Wall Receive}

Models the effects that occur when the periodic status poller, running
in the MCP server coroutine context, detects that the wall KV entry
has changed (typically after being woken by the NATS KV watcher).
The watcher itself runs \emph{outside} any tool handler context in a
background asyncio task and only wakes the poller; it does \emph{not}
update the wall description, modify the display queue, or fire an MCP
notification directly---sending notifications from a background task
is unreliable for the same reason as NATS subscription callbacks.

After being woken, the poller performs its next active tick (every
2~seconds), reads the wall value from KV, detects any change, applies
this schema to update the wall description and display queue, and then
fires the notification from the MCP server's coroutine context.

The precondition $natsConn = natsConnected$ implies
$session = sessCaptured$ (by state invariant), so the poller will
have a valid session when it fires the deferred notification.

\begin{schema}{KVWallReceive}
  \Delta NotificationState \\
  wallText? : MSG \\
  wallKey? : SOURCEKEY
  \where
  pendingAge' = pendingAge \\
  natsConn = natsConnected \\
  toolContext = zfalse \\
  wallDesc' \neq wallDesc \\
  wallDescChanged' = ztrue \\
  (wallKey? \notin \ran queueItems \implies \\
  \quad~ queueItems' = queueItems \cat \langle wallKey? \rangle \land \\
  \quad~ queueKinds' = queueKinds \cup \\
  \quad~ \quad~ \{ wallKey? \mapsto itemWall \} \land \\
  \quad~ queueTexts' = queueTexts \cup \\
  \quad~ \quad~ \{ wallKey? \mapsto wallText? \}) \\
  (wallKey? \in \ran queueItems \implies \\
  \quad~ queueItems' = queueItems \land \\
  \quad~ queueKinds' = queueKinds \land \\
  \quad~ queueTexts' = queueTexts) \\
  lastNotifyResult' = lastNotifyResult \\
  barStale' = barStale \\
  session' = session \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  talkDesc' = talkDesc \\
  talkDescChanged' = talkDescChanged \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{NATS Talk Callback}

Fires when a core NATS message arrives on the talk notification
subject.  Runs \emph{outside} any tool handler context in a NATS
subscription callback coroutine.  Mutates the talk description and
adds the talk item to the display queue, but does \emph{not} fire
an MCP notification directly---sending notifications from a NATS
callback is unreliable because Claude~Code intermittently ignores
them (different coroutine context from the MCP server).

Instead, the callback sets $talkDescChanged' = ztrue$ and the
background poller ($PollTick$) detects the change on its next tick
and fires the notification from the MCP server's own coroutine
context.  This is the same pattern wall uses: state mutation is
decoupled from notification delivery.

The precondition $talkSubscribed = ztrue$ implies
$natsConn = natsConnected$ (by state invariant), which in turn implies
$session = sessCaptured$.  The poller will therefore have a valid
session when it fires the deferred notification.

\begin{schema}{NatsTalkCallback}
  \Delta NotificationState \\
  senderBody? : MSG \\
  senderKey? : SOURCEKEY
  \where
  pendingAge' = pendingAge \\
  talkSubscribed = ztrue \\
  toolContext = zfalse \\
  talkBare = \emptyset \\
  talkQueued < maxQueueSize \\
  talkDesc' \neq talkDescDefault \\
  talkDescChanged' = ztrue \\
  queueItems' = queueItems \cat \langle senderKey? \rangle \\
  queueKinds' = queueKinds \oplus \\
  \quad~ \{ senderKey? \mapsto itemTalk \} \\
  queueTexts' = queueTexts \oplus \\
  \quad~ \{ senderKey? \mapsto senderBody? \} \\
  currentIndex' = \# queueItems \\
  lastNotifyResult' = lastNotifyResult \\
  barStale' = barStale \\
  session' = session \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued + 1 \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  wallDescChanged' = wallDescChanged \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Sync Unread File}

Writes the display queue snapshot to the filesystem status file.
Called after any state change that affects what the status bar shows.

\begin{schema}{SyncUnreadFile}
  \Delta NotificationState \\
  count? : \nat
  \where
  pendingAge' = pendingAge \\
  count? \leq maxUnreadCount \\
  fileExists' = ztrue \\
  fileUnreadCount' = count? \\
  fileItems' = queueItems \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Poll Tick (Background Poller)}

The 2-second background poller tick.  Runs outside tool context.
If a description has changed, fires notification via the suspenders
path.  The state invariant guarantees $session = sessCaptured$
whenever NATS is connected, so the suspenders path always succeeds.
Clears the changed flags regardless.

\begin{schema}{PollTick}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  toolContext = zfalse \\
  ((wallDescChanged = ztrue \lor \\
  \quad~ talkDescChanged = ztrue) \implies \\
  \quad~ lastNotifyResult' = notifySent \land \\
  \quad~ barStale' = zfalse) \\
  ((wallDescChanged = zfalse \land \\
  \quad~ talkDescChanged = zfalse) \implies \\
  \quad~ lastNotifyResult' = lastNotifyResult \land \\
  \quad~ barStale' = barStale) \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = zfalse \\
  talkDescChanged' = zfalse \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Status Bar Refresh}

Claude~Code polls the status line command on its own schedule.
Reads the filesystem status file and updates the display.
This is the slow path---no MCP involvement.

\begin{schema}{StatusBarRefresh}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  (fileExists = ztrue \land \\
  \quad~ fileItems \neq \langle \rangle \implies \\
  \quad~ barLine2' \neq idleMarker) \\
  (fileExists = zfalse \implies
  barLine2' = idleMarker) \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

%%----------------------------------------------------------------------
\section{Display Queue Operations}
%%----------------------------------------------------------------------

\subsection{Advance If Due}

Rotation: wall items cycle (persistent), talk items are
ephemeral (removed after one display turn).  The $currentIndex + 1$
expression indexes into the 1-based sequence.

Talk items at the current position are removed (ephemeral).
Wall items cycle: the cursor advances, but the item stays in the
queue.  With a single wall item, the cursor stays at zero.

\begin{schema}{AdvanceTalk}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  queueItems \neq \langle \rangle \\
  queueKinds(queueItems(currentIndex + 1)) = itemTalk \\
  \# queueItems' = \# queueItems - 1 \\
  queueKinds' = \\
  \quad~ \{ queueItems(currentIndex + 1) \} \ndres queueKinds \\
  queueTexts' = \\
  \quad~ \{ queueItems(currentIndex + 1) \} \ndres queueTexts \\
  \ran queueItems' = \ran queueItems \setminus \\
  \quad~ \{ queueItems(currentIndex + 1) \} \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\begin{schema}{AdvanceWall}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  queueItems \neq \langle \rangle \\
  queueKinds(queueItems(currentIndex + 1)) = itemWall \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  (\# queueItems > 1 \implies \\
  \quad~ currentIndex' = (currentIndex + 1) \mod \\
  \quad~ \quad~ \# queueItems) \\
  (\# queueItems = 1 \implies
  currentIndex' = 0) \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  readDesc' = readDesc \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

%%----------------------------------------------------------------------
\section{Tool-Description Marker Channels}
\label{sec:marker}
%%----------------------------------------------------------------------

The MCP server surfaces incoming activity by mutating a tool's
\emph{description} and firing \texttt{tools/list\_changed}: the agent
notices the changed text on its next turn.  Two such channels are
modelled here.  The \texttt{read\_messages} channel is the reference: it
is known to behave correctly.  The \texttt{talk} channel is the one under
examination.

The governing property, developed in Section~\ref{sec:markerprops}, is
that for each channel the marker is present exactly when the underlying
activity exists:
\[
  desc \neq default \iff activity.
\]
This section gives the operations of both channels.  It deliberately
gives them first \emph{without} the reconciling transitions that make the
property hold, so that Section~\ref{sec:markerprops} can exhibit the
defect as a reachable state before Section~\ref{sec:markerfix} repairs
it.

\subsection{Read Channel (the self-clearing reference)}

Mail arrives and raises the unread count; the description gains its
marker.  Reading mail returns the count to zero and the description
reverts.  Because the description is driven by a count that
\emph{decreases}, this channel clears itself: there is no way to leave
the marker set once the count reaches zero.

\begin{schema}{MailArrive}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  unreadCount < maxUnreadCount \\
  unreadCount' = unreadCount + 1 \\
  readDesc' \neq readDescDefault \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\begin{schema}{ReadMail}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  unreadCount > 0 \\
  unreadCount' = 0 \\
  readDesc' = readDescDefault \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Talk Channel (invite, message, drain)}

An invite arrives and is recorded in $talkPending$, retaining the
inviter's session key; the description gains its \texttt{[TALK]} marker.
A talk message arrives and raises $talkQueued$.  Both set the marker.

\begin{schema}{TalkInviteArrive}
  \Delta NotificationState \\
  inviter? : USER \\
  inviterKey? : SESSIONKEY
  \where
  pendingAge' = pendingAge \\
  natsConn = natsConnected \\
  inviter? \notin talkBare \\
  talkPending' = talkPending \oplus \{ inviter? \mapsto inviterKey? \} \\
  talkDesc' \neq talkDescDefault \\
  talkDescChanged' = ztrue \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\begin{schema}{TalkMessageArrive}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  natsConn = natsConnected \\
  talkQueued < maxQueueSize \\
  talkQueued' = talkQueued + 1 \\
  talkDesc' \neq talkDescDefault \\
  talkDescChanged' = ztrue \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

The defect (biff-9la, observed live).  The drain that ends a talk---the
aggregate of accepting, exchanging, and hanging up---empties every source
of talk activity, but leaves $talkDesc$ untouched.  Nothing reverts the
marker; there is no withdrawal frame and no expiry.  $TalkDrainStale$
models exactly this: it clears $talkPending$, $talkBare$, $talkQueued$
and $talkConnected$, yet preserves $talkDesc$.

\begin{schema}{TalkDrainStale}
  \Delta NotificationState
  \where
  pendingAge' = 0 \\
  (\dom talkPending \neq \emptyset \lor talkBare \neq \emptyset \lor
  talkQueued > 0 \lor talkConnected = ztrue) \\
  talkPending' = \emptyset \\
  talkBare' = \emptyset \\
  talkQueued' = 0 \\
  talkConnected' = zfalse \\
  talkDesc' = talkDesc \\
  talkDescChanged' = talkDescChanged \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

The coupled defect.  A second, independent fault records an invite
\emph{without} its session key---the tty is dropped on the way into the
pending representation.  $TalkInviteArriveBare$ puts the inviter into
$talkBare$ and raises the marker; the resulting accept hint reads a bare
\texttt{talk @user}, which fails at the prompt (``Talk needs a specific
session'').

\begin{schema}{TalkInviteArriveBare}
  \Delta NotificationState \\
  inviter? : USER
  \where
  pendingAge' = pendingAge \\
  natsConn = natsConnected \\
  inviter? \notin \dom talkPending \\
  talkBare' = talkBare \cup \{ inviter? \} \\
  talkDesc' \neq talkDescDefault \\
  talkDescChanged' = ztrue \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

%%----------------------------------------------------------------------
\section{Marker Consistency: the property and the defect}
\label{sec:markerprops}
%%----------------------------------------------------------------------

Write $talkActivity$ for the disjunction of the three talk activity
sources and $readActivity$ for a positive unread count:
\begin{eqnarray*}
  talkActivity & \defs & \dom talkPending \neq \emptyset \lor
    talkBare \neq \emptyset \lor talkQueued > 0 \lor
    talkConnected = ztrue \\
  readActivity & \defs & unreadCount > 0.
\end{eqnarray*}

The intended invariant is that each marker is present exactly when its
activity exists, and that a talk marker always carries a session-naming
accept hint:

\begin{schema}{MarkerConsistent}
  NotificationState
  \where
  (talkDesc \neq talkDescDefault) \iff \\
  \quad (\dom talkPending \neq \emptyset \lor talkBare \neq \emptyset \lor
  talkQueued > 0 \lor talkConnected = ztrue) \\
  (readDesc \neq readDescDefault) \iff unreadCount > 0
\end{schema}

\begin{schema}{HintNamesSession}
  NotificationState
  \where
  talkDesc \neq talkDescDefault \implies talkBare = \emptyset
\end{schema}

Neither property is yet part of $NotificationState$, so both may fail.
Two model-checking runs at set size~2 exhibit the failures as reachable
states.

\paragraph{The stale marker.}  Setting the ProB goal to
\[
  talkDesc \neq talkDescDefault \land \lnot\, talkActivity
\]
finds a reachable state: an invite arrives ($TalkInviteArrive$ or
$TalkMessageArrive$ raises the marker), then $TalkDrainStale$ empties
every activity source while preserving $talkDesc$.  The marker persists
with nothing behind it.  This is the biff-9la defect.

\paragraph{The bare hint.}  Setting the goal to
\[
  talkDesc \neq talkDescDefault \land talkBare \neq \emptyset
\]
finds a reachable state reached by $TalkInviteArriveBare$: the marker is
present but the pending invite carries no session key, so its accept hint
names no session.

\paragraph{Liveness.}  The safety goals above show a \emph{bad state} is
reachable.  The sharper statement is a liveness one: from some reachable
marker-present state, the marker can never clear.  Expressed as the CTL
formula
\[
  \mathrm{AG}\,(talkDesc \neq talkDescDefault \implies
    \mathrm{EF}\;talkDesc = talkDescDefault),
\]
the model refutes it, because $TalkDrainStale$ is the only way to empty
activity and it does not touch $talkDesc$: once set, the marker is
permanent.  The liveness refutation is robust---it holds even though
other operations remain enabled, so it is not masked the way a plain
deadlock check would be.

%%----------------------------------------------------------------------
\section{The Fix: reconciling transitions}
\label{sec:markerfix}
%%----------------------------------------------------------------------

The repair has two parts.  First, $MarkerConsistent$, $HintNamesSession$
and the read biconditional are promoted into the state schema
(Section~\ref{sec:state}), so they are now invariants that every
operation must respect.  Second, transitions are added that remove
activity \emph{and reconcile the description in the same step}.

Promoting the invariants has an immediate and welcome consequence: the
two defect operations become infeasible.  $TalkDrainStale$ empties
activity while preserving $talkDesc \neq talkDescDefault$; its post-state
now contradicts $MarkerConsistent$, so it is enabled in no reachable
state.  $TalkInviteArriveBare$ sets $talkBare \neq \emptyset$ while
raising the marker; its post-state contradicts $HintNamesSession$, so it
too is dead.  ProB reports both as uncovered operations.  This is the
point of the exercise: once the consistency of marker and activity is an
invariant, the specification \emph{proves} that no drain can strand the
marker and no invite can be recorded without its session key.  The defect
operations are kept in the text only as the negative results they now
are.

A dead operation is not a live transition, however.  Something must
still remove activity, and---because the invariant forbids emptying
activity without reverting the description---any such transition must do
both at once.  Four suffice: $TalkDrain$, $WithdrawArrive$,
$ExpirePendingInvite$ and $TalkAccept$; the proof in
Section~\ref{sec:markerfix2} shows each discharges an obligation the
others do not.  $WithdrawArrive$ and $ExpirePendingInvite$ together
replace the single abstract $DropPendingInvite$ of the earlier drafts:
the two concrete removal paths the implementation must supply---a
withdrawal frame and a time-to-live expiry---have distinct preconditions
and so are kept distinct in the model.

\subsection{Draining messages and ending the conversation}

$TalkDrain$ empties the message queue and ends the connection.  If no
pending invite remains behind, it reverts the description; otherwise the
marker stays, now backed only by the surviving invites.

\begin{schema}{TalkDrain}
  \Delta NotificationState
  \where
  pendingAge' = pendingAge \\
  (talkQueued > 0 \lor talkConnected = ztrue) \\
  talkQueued' = 0 \\
  talkConnected' = zfalse \\
  ((\dom talkPending = \emptyset \land talkBare = \emptyset) \implies \\
  \quad (talkDesc' = talkDescDefault \land talkDescChanged' = zfalse)) \\
  ((\dom talkPending \neq \emptyset \lor talkBare \neq \emptyset) \implies \\
  \quad (talkDesc' = talkDesc \land talkDescChanged' = talkDescChanged)) \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Withdrawing a pending invite (ntWithdraw frame)}

The earlier drafts collapsed the two ways a pending invite can be removed
into a single abstract operation.  The two mechanisms have genuinely
different preconditions, however, and the implementation must provide both,
so they are modelled here as distinct operations---the partition of
Section~\ref{sec:markerfix2} then owes a separate test for each.

$WithdrawArrive$ is the first: the recipient receives an \emph{ntWithdraw}
notification from the inviter---the inviter ran \texttt{talk\_end}, hung up,
or otherwise cancelled---and removes that inviter's pending entry.  A
withdrawal frame carries the inviter's originating session key in the field
$withdrawKey?$; this is the \emph{ntWithdraw} analogue of the $nfromKey$ that
$talk.tex$ guards on for accepts.  The removal is honoured only when that key
equals the session key stored against the pending invite---%
$inviter? \in \dom talkPending$ and $talkPending~inviter? = withdrawKey?$.
The guard is not decoration.  The talk transport is core NATS publish and
subscribe; it gives no cross-session ordering, so a withdrawal emitted by an
earlier session of the same user may be delivered \emph{after} a fresh invite
from a later session of that user.  A user invites us from session~$A$,
withdraws, and re-invites from session~$B$; because $talkPending$ is a
function keyed on the user, the re-invite supersedes the first, so
$talkPending$ now maps that user to $B$'s key.  A withdrawal keyed to
session~$A$ that arrives late must not disturb the live session~$B$
invite---and with the key guard it does not, because the stored key (session
$B$'s) differs from the withdrawal's key (session $A$'s), so the equality
$talkPending~inviter? = withdrawKey?$ fails and no removal occurs.  Without
the guard, the withdrawal would delete the current invite on a bare user
match, the marker would clear, and the live invite would be silently lost.  This is
exactly the stale-marker defect class the specification exists to prevent,
reached here by a non-adversarial reordering rather than by an attacker; the
key guard is the withdrawal-side counterpart of the accept-side consent
boundary of $talk.tex$ (DES-043, invariant~11).  A withdrawal whose key does
not match is not applied here; it is handled by $WithdrawForeign$ below.
$WithdrawArrive$'s precondition is otherwise the arrival of the frame on a
live connection; it does \emph{not} depend on how long the invite has been
pending.  If no activity survives, the description reverts and the clock
returns to zero.

\begin{schema}{WithdrawArrive}
  \Delta NotificationState \\
  inviter? : USER \\
  withdrawKey? : SESSIONKEY
  \where
  natsConn = natsConnected \\
  inviter? \in \dom talkPending \\
  talkPending~inviter? = withdrawKey? \\
  talkPending' = \{ inviter? \} \ndres talkPending \\
  talkBare' = talkBare \\
  ((\dom talkPending' = \emptyset \land talkBare' = \emptyset) \implies
  pendingAge' = 0) \\
  ((\dom talkPending' \neq \emptyset \lor talkBare' \neq \emptyset) \implies
  pendingAge' = pendingAge) \\
  ((\dom talkPending' = \emptyset \land talkBare' = \emptyset \land \\
  \quad talkQueued = 0 \land talkConnected = zfalse) \implies \\
  \quad (talkDesc' = talkDescDefault \land talkDescChanged' = zfalse)) \\
  ((\dom talkPending' \neq \emptyset \lor talkBare' \neq \emptyset \lor \\
  \quad talkQueued > 0 \lor talkConnected = ztrue) \implies \\
  \quad (talkDesc' = talkDesc \land talkDescChanged' = talkDescChanged)) \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{A withdrawal that names the wrong session (WithdrawForeign)}

$WithdrawForeign$ is the withdrawal-side counterpart of $talk.tex$'s
$DrainForeignAccept$.  A withdrawal frame arrives whose key does \emph{not}
match the pending invite it purports to cancel---either the user has no
pending invite at all, or the stored invite carries a different session key,
as in the reordering above.  The frame is dropped: the pending set, the
marker, the clock, and every other component are left exactly as they were.
Modelling it as a distinct operation, rather than as a silent no-op, forces
the test-partition of Section~\ref{sec:markerfix2} to owe a case for it, and
records in the specification that the implementation must recognise and
discard such a frame rather than acting on its user field alone.

\begin{schema}{WithdrawForeign}
  \Delta NotificationState \\
  inviter? : USER \\
  withdrawKey? : SESSIONKEY
  \where
  natsConn = natsConnected \\
  \lnot\, (inviter? \in \dom talkPending \land
  talkPending~inviter? = withdrawKey?) \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  pendingAge' = pendingAge \\
  talkDesc' = talkDesc \\
  talkDescChanged' = talkDescChanged \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{The withdrawal gap, as a dead operation (WithdrawStale)}

$WithdrawStale$ is the defect this section exists to close, kept in the text
as the negative result it now is.  It models the username-only withdrawal:
an \emph{ntWithdraw} frame arrives whose key does not match the stored invite
($talkPending~inviter? \neq withdrawKey?$), yet it removes the invite anyway,
on the bare user match.  The removed invite is the live one; the marker
clears with nothing behind it and the current invite is silently lost.  The
history component $staleWithdrawHit$ records that this has happened, so the
suppression can be named as a reachable state.

Promoting $staleWithdrawHit = zfalse$ into $NotificationState$
(Section~\ref{sec:state}) makes $WithdrawStale$ infeasible: its post-state
sets $staleWithdrawHit' = ztrue$, which the invariant forbids.  ProB reports
it as an uncovered operation, and the goal $staleWithdrawHit = ztrue$ is
unreachable.  Delete the invariant---as an earlier, username-only draft
implicitly did---and $WithdrawStale$ becomes enabled: with two distinct
session keys it drives the model to $staleWithdrawHit = ztrue$, the reordering
counterexample of Section~\ref{sec:markerfix2}.  This is the same device by
which $TalkDrainStale$ and $TalkInviteArriveBare$ are proved dead against
$MarkerConsistent$ and $HintNamesSession$.

\begin{schema}{WithdrawStale}
  \Delta NotificationState \\
  inviter? : USER \\
  withdrawKey? : SESSIONKEY
  \where
  natsConn = natsConnected \\
  inviter? \in \dom talkPending \\
  talkPending~inviter? \neq withdrawKey? \\
  talkPending' = \{ inviter? \} \ndres talkPending \\
  talkBare' = talkBare \\
  staleWithdrawHit' = ztrue \\
  ((\dom talkPending' = \emptyset \land talkBare' = \emptyset) \implies
  pendingAge' = 0) \\
  ((\dom talkPending' \neq \emptyset \lor talkBare' \neq \emptyset) \implies
  pendingAge' = pendingAge) \\
  ((\dom talkPending' = \emptyset \land talkBare' = \emptyset \land \\
  \quad talkQueued = 0 \land talkConnected = zfalse) \implies \\
  \quad (talkDesc' = talkDescDefault \land talkDescChanged' = zfalse)) \\
  ((\dom talkPending' \neq \emptyset \lor talkBare' \neq \emptyset \lor \\
  \quad talkQueued > 0 \lor talkConnected = ztrue) \implies \\
  \quad (talkDesc' = talkDesc \land talkDescChanged' = talkDescChanged)) \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult
\end{schema}

\subsection{Expiring a pending invite (time-to-live)}

The second mechanism covers the inviter who never sends a withdrawal---a
crash, a dropped connection, an abandoned session.  No frame arrives, so
$WithdrawArrive$ is never enabled for that invite; the only thing that
changes is the clock.  $AgeTick$ models the passage of time on a poller
tick: while any invite is pending and the clock is below its bound, the
clock advances by one.  It touches nothing else---not the pending set, not
the description, not the activity.

\begin{schema}{AgeTick}
  \Delta NotificationState
  \where
  toolContext = zfalse \\
  (\dom talkPending \neq \emptyset \lor talkBare \neq \emptyset) \\
  pendingAge < maxInviteAge \\
  pendingAge' = pendingAge + 1 \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  talkDesc' = talkDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  talkDescChanged' = talkDescChanged \\
  unreadCount' = unreadCount \\
  talkPending' = talkPending \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

$ExpirePendingInvite$ then reaps an invite once its clock has reached the
bound.  Its precondition---$pendingAge = maxInviteAge$ on a poller
tick---is distinct from $WithdrawArrive$'s, which fires on frame receipt at
any age.  The state effect is the same: the inviter's entry is removed, and
if no activity survives, the description reverts and the clock returns to
zero.

\begin{schema}{ExpirePendingInvite}
  \Delta NotificationState \\
  inviter? : USER
  \where
  toolContext = zfalse \\
  pendingAge = maxInviteAge \\
  (inviter? \in \dom talkPending \lor inviter? \in talkBare) \\
  talkPending' = \{ inviter? \} \ndres talkPending \\
  talkBare' = talkBare \setminus \{ inviter? \} \\
  ((\dom talkPending' = \emptyset \land talkBare' = \emptyset) \implies
  pendingAge' = 0) \\
  ((\dom talkPending' \neq \emptyset \lor talkBare' \neq \emptyset) \implies
  pendingAge' = pendingAge) \\
  ((\dom talkPending' = \emptyset \land talkBare' = \emptyset \land \\
  \quad talkQueued = 0 \land talkConnected = zfalse) \implies \\
  \quad (talkDesc' = talkDescDefault \land talkDescChanged' = zfalse)) \\
  ((\dom talkPending' \neq \emptyset \lor talkBare' \neq \emptyset \lor \\
  \quad talkQueued > 0 \lor talkConnected = ztrue) \implies \\
  \quad (talkDesc' = talkDesc \land talkDescChanged' = talkDescChanged)) \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  talkQueued' = talkQueued \\
  talkConnected' = talkConnected \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

\subsection{Accepting an invite}

$TalkAccept$ is the ordinary path: a pending invite is consumed and the
conversation becomes live.  Activity does not vanish---it moves from
$talkPending$ to $talkConnected$---so the marker is preserved, consistent
with the invariant.

\begin{schema}{TalkAccept}
  \Delta NotificationState \\
  inviter? : USER
  \where
  inviter? \in \dom talkPending \\
  talkPending' = \{ inviter? \} \ndres talkPending \\
  talkConnected' = ztrue \\
  ((\dom talkPending' = \emptyset \land talkBare' = \emptyset) \implies
  pendingAge' = 0) \\
  ((\dom talkPending' \neq \emptyset \lor talkBare' \neq \emptyset) \implies
  pendingAge' = pendingAge) \\
  talkDesc' = talkDesc \\
  talkDescChanged' = talkDescChanged \\
  natsConn' = natsConn \\
  kvProvisioned' = kvProvisioned \\
  talkSubscribed' = talkSubscribed \\
  session' = session \\
  toolContext' = toolContext \\
  wallDesc' = wallDesc \\
  readDesc' = readDesc \\
  wallDescChanged' = wallDescChanged \\
  unreadCount' = unreadCount \\
  talkBare' = talkBare \\
  talkQueued' = talkQueued \\
  queueItems' = queueItems \\
  queueKinds' = queueKinds \\
  queueTexts' = queueTexts \\
  currentIndex' = currentIndex \\
  fileExists' = fileExists \\
  fileUnreadCount' = fileUnreadCount \\
  fileItems' = fileItems \\
  barLine2' = barLine2 \\
  barStale' = barStale \\
  lastNotifyResult' = lastNotifyResult \\
  staleWithdrawHit' = staleWithdrawHit
\end{schema}

%%----------------------------------------------------------------------
\section{Marker Consistency: the proof}
\label{sec:markerfix2}
%%----------------------------------------------------------------------

With the invariants promoted and the three reconciling transitions in
place, three checks at set size~2 discharge the obligations.

\paragraph{Safety.}  Ordinary model checking visits every reachable state
and finds no invariant violation and no deadlock.  Because
$MarkerConsistent$ and $HintNamesSession$ are now part of the state
schema, this establishes that no reachable state has a marker without its
activity, and none has a talk marker whose hint names no session.  The
two goal predicates that were reachable in
Section~\ref{sec:markerprops}---$talkDesc \neq talkDescDefault \land
\lnot\, talkActivity$ and $talkDesc \neq talkDescDefault \land talkBare
\neq \emptyset$---are now unreachable.

\paragraph{Liveness.}  The CTL formula
$\mathrm{AG}\,(talkDesc \neq talkDescDefault \implies
\mathrm{EF}\; talkDesc = talkDescDefault)$
now holds: from every marker-present state at least one reconciling
transition is enabled and leads, directly or through $TalkAccept$, to a
state in which the marker has cleared.

\paragraph{Necessity of each transition.}  Each reconciling transition
discharges a distinct obligation.  $TalkDrain$ is the only transition
that clears $talkQueued$ or $talkConnected$; without it a marker backed
by a queued message or a live connection can never clear.
$WithdrawArrive$ and $ExpirePendingInvite$ are the two transitions that
remove a pending invite without the recipient's acceptance, and they are
enabled under different conditions: $WithdrawArrive$ on receipt of an
\emph{ntWithdraw} frame from an inviter who cancels, $ExpirePendingInvite$
on a poller tick once the clock has reached $maxInviteAge$.  The second is
what covers exactly the biff-9la scenario---an invite from a peer who
never returns and never withdraws: only the time-to-live path clears it,
so without $ExpirePendingInvite$ (and the $AgeTick$ that drives its
clock) the talk channel would not be self-clearing in the way the read
reference is.  $TalkAccept$ is the ordinary progress transition; it is not
required for the marker to clear, but without it the only exit from a
pending invite is to remove it, and the handshake modelled in $talk.tex$
could never complete.

%%----------------------------------------------------------------------
\section{System Properties}
%%----------------------------------------------------------------------

\subsection{Notification Delivery Guarantee}

The safety property: whenever a tool description changes,
Claude~Code is notified.  In the corrected system this holds
by construction---no operation produces $notifyDropped$ after
a description change.

\begin{schema}{NotifyGuarantee}
  NotificationState
  \where
  (wallDescChanged = ztrue \lor talkDescChanged = ztrue)
  \implies lastNotifyResult = notifySent
\end{schema}

\emph{Proof sketch}: $WallPost$ sets $wallDescChanged' = ztrue$ and
$lastNotifyResult' = notifySent$ (belt path, tool context).
$KVWallReceive$ and $NatsTalkCallback$ set their respective changed
flags but defer notification to the poller: they preserve
$lastNotifyResult$, and $PollTick$ detects the changed flags and sets
$lastNotifyResult' = notifySent$.  The invariant holds because,
whenever a description change sets a ``changed'' flag, the next execution
of $PollTick$ that observes this flag produces $notifySent$; thus no
changed flag can persist across two consecutive $PollTick$ executions
without producing $notifySent$.  Both KV watchers and NATS callbacks
wake the poller from napping, ensuring $PollTick$ runs promptly even
during idle periods.

\subsection{Session Invariant}

The key structural invariant: NATS connectivity implies
session capture.

\begin{schema}{SessionInvariant}
  NotificationState
  \where
  natsConn = natsConnected \implies session = sessCaptured
\end{schema}

This holds because:
\begin{enumerate}
  \item $Init$ sets $natsConn = natsDisconnected$ (antecedent false).
  \item $CaptureSession$ sets $session = sessCaptured$ while
    $natsConn$ remains $natsDisconnected$ (antecedent false).
  \item $EnsureConnected$ sets $natsConn = natsConnected$ and
    preserves $session$. The invariant in $NotificationState$
    forces $session = sessCaptured$ as an implicit precondition.
  \item No operation transitions $session$ from $sessCaptured$
    back to $sessNone$.
  \item $CloseConnection$ sets $natsConn = natsClosed$ (antecedent false).
\end{enumerate}

\subsection{Suspenders Always Available}

When NATS is connected and no tool context is active, the
suspenders path is always applicable:

\begin{schema}{SuspendersAvailable}
  NotificationState
  \where
  natsConn = natsConnected \\
  toolContext = zfalse \\
  session = sessCaptured
\end{schema}

The last conjunct is not an additional requirement---it follows from
the first conjunct via $SessionInvariant$.  This means every poller
tick with pending changes can successfully send a notification.
Neither NATS callbacks nor KV watchers send notifications directly;
they set flags that the poller consumes.

\subsection{Wall and Talk Symmetry}

In the corrected system, both wall and talk notifications always
reach Claude~Code.  Wall posting uses the belt path (tool context);
wall reception from another user uses $KVWallReceive$ which defers to
the poller.  Talk reception uses $NatsTalkCallback$ which also defers
to the poller.  Both ultimately produce $lastNotifyResult = notifySent$.

\paragraph{Erratum (biff-9la).}  The symmetry stated above concerns only
\emph{notification delivery}---that a \texttt{tools/list\_changed} is
sent.  It does not extend to the \emph{life-cycle of the tool
description}, and the earlier drafts of this specification wrongly implied
that it did by modelling $talkDesc$ as an opaque value with no invariant.
Wall and talk are not symmetric here.  The wall and read markers are
driven by sources that decrease---an unread count that falls to zero, a
display item that is advanced away---so they clear themselves.  The talk
marker is driven by $talkPending$, $talkQueued$ and $talkConnected$, and
nothing in the original model reverted $talkDesc$ when those emptied:
Sections~\ref{sec:markerprops} reproduces the resulting stale marker as a
reachable state.  The corrected treatment requires the reconciling
transitions of Section~\ref{sec:markerfix}; only then does the talk
description track its activity as the read reference does.  The
consistency invariant $MarkerConsistent$, not delivery symmetry, is the
property that matters.

\paragraph{Erratum ($talk.tex$ invariant~4).}  Invariant~4 of $talk.tex$
(``Pending invites \ldots remain in the set until the next \texttt{talk}
command'') describes exactly the grow-only behaviour that strands the
marker.  If pending invites are never removed except by a subsequent
\texttt{talk} command, then a marker driven by $talkPending$ can never
clear on its own.  Under the invariant $MarkerConsistent$ that reading is
superseded: a pending invite must also be removable by $WithdrawArrive$---an
\emph{ntWithdraw} frame from the inviter---or by $ExpirePendingInvite$---a
time-to-live expiry aged out on a poller tick---so that the talk channel
is self-clearing.  The consent-boundary guarantees of $talk.tex$
(invariant~11) are unaffected: withdrawal and expiry only \emph{remove}
pending invites, never complete a handshake, so they cannot forge
consent.

\begin{schema}{WallNotifySuccess}
  NotificationState
  \where
  wallDescChanged = ztrue \\
  lastNotifyResult = notifySent
\end{schema}

\begin{schema}{TalkNotifySuccess}
  NotificationState
  \where
  talkDescChanged = ztrue \\
  lastNotifyResult = notifySent
\end{schema}

\subsection{Startup Sequence}

The correct startup sequence is:
\begin{enumerate}
  \item $Init$ --- all state zeroed, NATS disconnected, session uncaptured.
  \item $CaptureSession$ --- session captured from server lifespan context.
  \item $EnsureConnected$ --- NATS connected (invariant satisfied).
  \item Poller starts, subscriptions begin --- suspenders path available.
\end{enumerate}

Any reordering that attempts $EnsureConnected$ before $CaptureSession$
would violate the state invariant and is therefore rejected by the
specification.

\end{document}
